---
output: pdf_document
---

\centering
\Huge

**Big Data Cup 2021 - Data Preparation**

\raggedright

\rule{\textwidth}{1.5pt}
\small

```{r setup, include = FALSE}
library(knitr) # For formatting the document
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

Summary: This file includes the data wrangling process to trim the plays down to entry-to-exit sequences. 


# Load Packages and Datasets

Load necessary packages

```{r load packages}
library(tidyverse) # For various data analysis functions

library(furrr) # For parallelizing nested data computation

library(here) # To call or save code/data in certain folders

library(reshape2) # For the melt() and dcast() functions

library(sf) # For spatial data analysis

library(zoo) # For na.locf() function

source(here::here("Code/0. Functions/Plotting-Functions.R"))
```

Load data sets

```{r load data}
otters_data = read_csv("https://raw.githubusercontent.com/bigdatacup/Big-Data-Cup-2021/main/hackathon_scouting.csv")
```

Parallelize code

```{r parallel}
# plan(multicore) # Uncomment if you are on a Mac to run code in parallel
# plan(multisession) # Uncomment if you are on a Windows to run code in parallel
```

# Extracting Plays Meeting Model Specifications (Zone Entry to End of Possession)

We will be focusing on offensive play following a zone entry for our project. This code chunk breaks the play down into sequences of offensive zone play beginning with an entry and ending with an exit or a whistle.

```{r entry to exit}
## STEP 1: BASIC VARIABLE CLEAN UP ##
otters_data_clean_v1 = otters_data %>%
  # Rename variables to remove spaces
  rename(x_coord = `X Coordinate`, y_coord = `Y Coordinate`, x_coord_2 = `X Coordinate 2`, y_coord_2 = `Y Coordinate 2`,
         Player_2 = `Player 2`, Detail_1 = `Detail 1`, Detail_2 = `Detail 2`, Detail_3 = `Detail 3`, Detail_4 = `Detail 4`,
         Home_Team = `Home Team`, Away_Team = `Away Team`, Home_Team_Goals = `Home Team Goals`, Away_Team_Goals = `Away Team Goals`,
         Home_Team_Skaters = `Home Team Skaters`, Away_Team_Skaters = `Away Team Skaters`) %>%
  # Create a variable "Total Seconds" that treats time as a numeric value
  separate(Clock, into = c("Minutes", "Seconds"), sep = ":", remove = FALSE) %>%
  mutate(Minutes = as.numeric(Minutes), Seconds = as.numeric(Seconds)) %>%
  mutate(Total_Seconds = 1200 - (Minutes * 60 + Seconds)) %>%
  # Remove penalties
  filter(Event != "Penalty Taken")


## STEP 2: BREAK DATA DOWN INTO WHISTLE-TO-WHISTLE "EPISODES" AND THEN FURTHER BREAK IT DOWN INTO ENTRY-TO-ENTRY PLAYS ##
otters_data_clean_v2 = otters_data_clean_v1 %>%
  # Create a binary variable to mark faceoffs
  mutate(faceoff = ifelse(Event == "Faceoff Win", 1, 0)) %>%
  # Also mark last event before the faceoff with a binary variable
  mutate(last_event_b4_faceoff = lead(faceoff)) %>%
  # Create an episode ID that increases by one at each faceoff
  mutate(episode_id = cumsum(faceoff)) %>%
  # Nest the data by episode_id
  group_by(episode_id) %>%
  nest() %>%
  # For each episode...
  mutate(data = future_map(data, ~.x %>% 
                      # Create a binary variable to mark zone entries
                      mutate(entry = ifelse(Event == "Zone Entry", 1, 0)) %>%
                      # Create an entry count to associate each play with an entry
                      mutate(entry_count = cumsum(entry)) %>%
                      # Remove all plays before an entry (faceoff to first zone entry)
                      filter(entry_count > 0))) %>%
  # Unnest the data
  unnest() %>%
  # Create an entry ID using the episode_id and the entry_count
  mutate(entry_id = episode_id * 100 + entry_count)


## STEP 3: GET BASIC INFO FOR ENTRY (TEAM, TIME, STRENGTH STATE, ETC.) ##
otters_data_clean_v3 = otters_data_clean_v2 %>%
  # Nest data by entry-to-entry play
  group_by(entry_id) %>%
  nest() %>%
  # For each entry-to-entry play...
  mutate(data = future_map(data, ~.x %>%
                      # Label the offensive team
                      mutate(entry_team = .x %>% filter(entry == 1) %>% select(Team) %>% unlist()) %>%
                      # Determine the entry time
                      mutate(entry_time = .x %>% filter(entry == 1) %>% select(Total_Seconds) %>% unlist()))) %>%
  # Unnest the data
  unnest() %>%
  # Find the time since entry for each play
  mutate(time_since_entry = Total_Seconds - entry_time) %>%
  # Determine how many skaters the offensive and defensive team have
  mutate(offensive_team_skaters = ifelse(entry_team == Home_Team, Home_Team_Skaters, Away_Team_Skaters),
         defensive_team_skaters = ifelse(entry_team == Home_Team, Away_Team_Skaters, Home_Team_Skaters)) %>%
  # Reorient xy-coordinates so that they are uniform for both teams (not always facing left)
  mutate(same_direction_x = ifelse(Team == entry_team, x_coord, 200 - x_coord),
         same_direction_x2 = ifelse(Team == entry_team, x_coord_2, 200 - x_coord_2),
         same_direction_y = ifelse(Team == entry_team, y_coord, 85 - y_coord),
         same_direction_y2 = ifelse(Team == entry_team, y_coord_2, 85 - y_coord_2))


## STEP 4: REMOVE ALL PLAYS OUTSIDE OF THE OFFENSIVE ZONE TO CONVERT EACH PLAY FROM ENTRY-TO-ENTRY TO ENTRY-TO-EXIT ##
otters_data_clean_v4 = otters_data_clean_v3 %>%
  # Filter out all plays that occur outside of offensive zone (except for entries)
  filter((same_direction_x >= 125) | Event == "Zone Entry")
  

## STEP 5: REMOVE ODDITIES ##
otters_data_clean_v5 = otters_data_clean_v4 %>%
  ## 5.1. Icings
  # Identify icing plays
  mutate(icing = ifelse((last_event_b4_faceoff == 1) & (entry == 1) & (x_coord <= 100), 1, 0)) %>%
  # Remove icings (Note: I checked and all icing plays are listed as their own play (ie unique entry ID), I left the check out for quicker computation)
  filter(icing == 0)


## STEP 6: REMOVE ANY UNNEEDED COLUMNS ##
otters_data_clean_v6 = otters_data_clean_v5 %>%
  # Remove binary variables that are no longer needed and minutes & seconds
  select(-faceoff, -icing, -Minutes, -Seconds)
```

Save this cleaned up data as a csv.

```{r save1}
write.csv(otters_data_clean_v6, here::here("Data/otters_entry_data.csv"), row.names = FALSE)
```



# Adding in Latent Movement for Visualization

For visualization purposes, we can add in "latent movements" between each observed play. 

```{r latent carries}
## ADD IN LATENT MOVEMENTS TO THE DATA ##
otters_latent_movement = otters_data_clean_v6 %>%
  # Nest the data by entry
  group_by(entry_id) %>%
  nest() %>%
  # Add in an additional row between each event
  mutate(data = future_map(data, ~.x %>%
                      # Use rownames as an index column and group by this index
                      rownames_to_column("event_id") %>%
                      group_by(event_id) %>%
                      # Add in an empty row between each group
                      do(rbind(., c(.$id, rep(NA, NCOL(.x) - 1)))) %>%
                      # Label the event of each new row as "Latent Movement"
                      mutate(Event = ifelse(Event %in% NA, "Latent Movement", Event)) %>%
                      # Ungroup the data
                      ungroup())) %>%
  # Add in the xy-coordinates for the latent movements
  mutate(data = future_map(data, ~.x %>%
                      # Add in x-coordinates for the latent movements
                      mutate(same_direction_x = case_when(
                        # If the previous event was a pass and the current event is latent, then the x-coordinate is the previous end x-coordinate
                        (Event == "Latent Movement") & (lag(Event) %in% c("Play", "Incomplete Play")) ~ lag(same_direction_x2),
                        # Else if the current event is 
                        Event == "Latent Movement" ~ lag(same_direction_x),
                        # Else leave it alone
                        TRUE ~ same_direction_x
                      )) %>%
                      # Add in y-coordinates for the latent movements
                      mutate(same_direction_y = case_when(
                        # Same idea as the x-coordinate
                        (Event == "Latent Movement" & lag(Event) %in% c("Play", "Incomplete Play")) ~ lag(same_direction_y2),
                        Event == "Latent Movement" ~ lag(same_direction_y),
                        TRUE ~ same_direction_y
                      )) %>%
                      # Add in the end xy-coordinates for latent movements
                      mutate(same_direction_x2 = ifelse(Event == "Latent Movement", lead(same_direction_x), same_direction_x2),
                             same_direction_y2 = ifelse(Event == "Latent Movement", lead(same_direction_y), same_direction_y2)) %>%
                      # Add in a variable to signify if an event is latent
                      mutate(is_latent = factor(ifelse(Event == "Latent Movement", "Latent", "Observed"), 
                                                levels = c("Observed", "Latent"))) %>%
                      # Remove the final latent movement of the play (not needed)
                      head(-1))) %>%
  # Unnest the data
  unnest() %>%
  # Remove latent movements that start and end in the same spot
  filter(!( (Event == "Latent Carry") & (same_direction_x == same_direction_x2) & (same_direction_y == same_direction_y2) )) %>%
  # Add in the associated team for each latent movement
  mutate(Team = ifelse(Event == "Latent Movement", lag(Team), Team))
```

Using ggplot we can visualize an entry-to-exit sequence

```{r}
## A SAMPLE PLAY VISUALIZATION ##
ggplot(otters_latent_movement %>% filter(entry_id == 401)) %>%
  fullrink_white() +
  geom_point(aes(x = same_direction_x, y = same_direction_y, fill = Event, alpha = is_latent), shape = 21, size = 3) +
  geom_segment(aes(x = same_direction_x, y = same_direction_y, xend = same_direction_x2, yend = same_direction_y2, 
                   colour = Team, linetype = is_latent), 
               arrow = arrow(length = unit(0.35,"cm"))) +
  scale_alpha_manual(values = c(1,0)) +
  scale_colour_manual(values = c("forestgreen", "purple")) +
  scale_linetype_manual(values = c("solid", "dashed")) +
  guides(alpha = "none") +
  labs(colour = "Movement Type", linetype = "Movement Type", fill = "Event Type")
```

Save the latent movement data as a csv.

```{r save2}
write.csv(otters_latent_movement, here::here("Data/otters_latent_movement.csv"), row.names = FALSE)
```



# Cleaning Data to Meet Markov Decision Process Assumptions

Below we go through the rigorous process of wrangling our data to fit the Markov Decision Process as described in our paper.

```{r data wrangling}
# Load in entry data
otters_data_clean_v6 = read_csv(here::here("Data/otters_entry_data.csv"))

## STEP 1: RELABEL SOME EVENTS (FOR NOW) ##
otters_data_clean_v7 = otters_data_clean_v6 %>%
  mutate(Event = ifelse(Event %in% c("Puck Recovery", "Takeaway"), "Recovery", Event)) %>%
  mutate(Event = ifelse(Event == "Play", "Pass", Event)) %>%
  mutate(Event = ifelse(Event == "Incomplete Play", "Failed Pass", Event))

## STEP 2: SPLIT SEQUENCES INTO SUBSEQUENCES OF CONSTANT TEAM POSSESSION ##
otters_data_clean_v8 = otters_data_clean_v7 %>%
  # Nest by entry_id
  group_by(entry_id) %>%
  nest() %>%
  # For each entry-to-exit play...
  mutate(data = future_map(data, ~.x %>% 
                      # Mark where there are changes in possession
                      mutate(poss_change = ifelse(Team != lag(Team), 1, 0)) %>%
                      # Label NAs as 0's
                      mutate(poss_change = ifelse(poss_change %in% NA, 0, poss_change)) %>%
                      # Use the possession change markers and cumsum to create a possession ID for each subsequence of team possessions
                      mutate(poss_id = cumsum(poss_change)) %>%
                      # Label the last play of each sequence
                      mutate(last_play = ifelse(lead(Team) %in% NA, 1, 0)))) %>%
  # Unnest the data
  unnest()

## STEP 3: DETERMINE WHETHER POSSESSION SUBSEQUENCES ARE OFFENSIVE AND DEFENSIVE ##
otters_data_clean_v9 = otters_data_clean_v8 %>%
  # Store the Event name as raw_event in case we need it
  mutate(raw_event = Event) %>%
  # Nest by entry and possession subsequence
  group_by(entry_id, poss_id) %>%
  nest() %>%
  # Determine the team with possession
  mutate(p_team = map(data, ~unique(.x$Team))) %>%
  # Determine the offensive team
  mutate(e_team = map(data, ~unique(.x$entry_team))) %>%
  # Determine if the team with possession is on offence
  mutate(on_offence = map2(p_team, e_team, ~.x == .y) %>% ifelse(1, 0))


## STEP 4: SIMPLIFY DEFENSIVE TEAM POSSESIONS AS JUST TURNOVERS AND EXITS
otters_data_clean_v10 = otters_data_clean_v9 %>%
  # Determine how many events are in each subsequence
  mutate(length = map(data, ~nrow(.x))) %>%
  # If the team with possession...
  mutate(data = case_when(
    # Isn't on offence and has >1 consecutive events...
    on_offence == 0 & length > 1 ~
      # Remove all but the first and last play
      future_map(data, ~rbind(head(.x, 1), tail(.x, 1)) %>%
            # Relabel first plays after puck acquisition by defensive team as turnovers
            mutate(Event = ifelse(Event %in% c("Pass", "Failed Pass", "Recovery"), "Turnover", Event)) %>%
            # Label last events as exits (except last events before the whistle, since those might not be exits)
            mutate(Event = ifelse(last_play == 1 & last_event_b4_faceoff == 0, "Exit", Event))),
    # Isn't on offence and has a single event...
    on_offence == 0 & length == 1 ~
      # Remove all but the first and last play
      future_map(data, ~.x %>%
            # Relabel first plays after puck acquisition by defensive team as turnovers
            mutate(Event = ifelse(Event %in% c("Pass", "Failed Pass", "Recovery"), "Turnover", Event)) %>%
            # Label last events as exits (except last events before the whistle, since those might not be exits)
            mutate(Event = ifelse(last_play == 1 & last_event_b4_faceoff == 0, "Exit", Event))),
    # Else, leave it
    TRUE ~ data
    )) %>%
  # For each subsequence determine if it is the last subseq in the full sequence or not
  mutate(is_last_play = map(data, ~sum(.x$last_play))) %>%
  # If the team with possession is on defence and the subsequence is not the exit play...
  mutate(data = ifelse(is_last_play == 0 & on_offence == 0,
                       # Just take first row (turnover)
                       map(data, ~head(.x, 1)),
                       # Else, leave it
                       data
                       )) %>%
  # Unnest and ungroup
  unnest() %>% ungroup()


  

## STEP 5: ADD IN RECOVERIES BETWEEN SHOT-SHOT EVENTS, RECOVERIES AFTER FAILED PASSES WHERE POSESSION IS MAINTAINED ##
##         AND TURNOVERS BEFORE EXITS BY DEFENSIVE TEAMS ##
otters_data_clean_v11 = otters_data_clean_v10 %>%
  # Create a binary variable to identify shot-shot (or shot-goal) plays
  mutate(shot_shot = ifelse((Event == "Shot") & (lag(Event) %in% c("Shot", "Goal")), 1, 0)) %>%
  # Create an count variable that adds 1 in between every shot-shot play
  mutate(temp_id = cumsum(shot_shot)) %>%
  # Group by this variable
  group_by(temp_id) %>%
  # Add in a row of NAs between each group
  do(rbind(., c(.$id, rep(NA, NCOL(.) - 1)))) %>%
  # Ungroup
  ungroup() %>%
  # Label every new row as a recovery
  mutate(Event = ifelse(Event %in% NA, "Recovery", Event)) %>%
  # Repeat the same process with recoveries after failed passes
  mutate(fpass = ifelse((Event %in% c("Turnover", "Exit", "Zone Entry")) & (lag(Event) == "Failed Pass"), 1, 0)) %>%
  mutate(temp_id = cumsum(fpass)) %>%
  group_by(temp_id) %>%
  do(rbind(., c(.$id, rep(NA, NCOL(.) - 1)))) %>%
  ungroup() %>%
  # Repeat the same process with turnovers between offensive team possession and defensive team exits
  mutate(nonto_exit = ifelse((Event == "Exit") & (lag(Event) != "Turnover"), 1, 0)) %>%
  mutate(temp_id = cumsum(nonto_exit)) %>%
  group_by(temp_id) %>%
  do(rbind(., c(.$id, rep(NA, NCOL(.) - 1)))) %>%
  ungroup() %>%
  # Repeat same process for turnover (def possession) to pass (off possession)
  mutate(to_pass = ifelse((Event %in% c("Pass", "Failed Pass")) & (lag(Event) == "Turnover"), 1, 0)) %>%
  mutate(temp_id = cumsum(to_pass)) %>%
  group_by(temp_id) %>%
  do(rbind(., c(.$id, rep(NA, NCOL(.) - 1)))) %>%
  ungroup() %>%
  # Fill in each NA value with the data in the following row (the xy-coordinates of our added events will be )
  na.locf(na.rm = FALSE, fromLast = TRUE)

# Recreate the same data frame but without filling in NA values with the following row's data
otters_data_clean_v11a = otters_data_clean_v10 %>%
  mutate(shot_shot = ifelse((Event == "Shot") & (lag(Event) == "Shot"), 1, 0)) %>%
  mutate(temp_id = cumsum(shot_shot)) %>%
  group_by(temp_id) %>%
  do(rbind(., c(.$id, rep(NA, NCOL(.) - 1)))) %>%
  ungroup() %>%
  mutate(Event = ifelse(Event %in% NA, "Recovery", Event)) %>%
  mutate(fpass = ifelse((Event %in% c("Turnover", "Exit", "Zone Entry")) & (lag(Event) == "Failed Pass"), 1, 0)) %>%
  mutate(temp_id = cumsum(fpass)) %>%
  group_by(temp_id) %>%
  do(rbind(., c(.$id, rep(NA, NCOL(.) - 1)))) %>%
  ungroup() %>%
  mutate(Event = ifelse(Event %in% NA, "Recovery", Event)) %>%
  mutate(nonto_exit = ifelse((Event == "Exit") & (lag(Event) != "Turnover"), 1, 0)) %>%
  mutate(temp_id = cumsum(nonto_exit)) %>%
  group_by(temp_id) %>%
  do(rbind(., c(.$id, rep(NA, NCOL(.) - 1)))) %>%
  ungroup() %>%
  mutate(Event = ifelse(Event %in% NA, "Turnover", Event)) %>%
  mutate(to_pass = ifelse((Event %in% c("Pass", "Failed Pass")) & (lag(Event) == "Turnover"), 1, 0)) %>%
  mutate(temp_id = cumsum(to_pass)) %>%
  group_by(temp_id) %>%
  do(rbind(., c(.$id, rep(NA, NCOL(.) - 1)))) %>%
  ungroup() %>%
  mutate(Event = ifelse(Event %in% NA, "Recovery", Event))

# Remove a few NA columns that arise
otters_data_clean_v11 = otters_data_clean_v11[1:36650,]
otters_data_clean_v11a = otters_data_clean_v11a[1:36650,]

# For columns that we don't want to fill in with the following row's data, use otters_data_clean_v11a
otters_data_clean_v11$Event = otters_data_clean_v11a$Event
otters_data_clean_v11$Player_2 = otters_data_clean_v11a$Player_2
otters_data_clean_v11$same_direction_x2 = otters_data_clean_v11a$same_direction_x2
otters_data_clean_v11$same_direction_y2 = otters_data_clean_v11a$same_direction_y2
otters_data_clean_v11$Detail_1 = otters_data_clean_v11a$Detail_1
otters_data_clean_v11$Detail_2 = otters_data_clean_v11a$Detail_2
otters_data_clean_v11$Detail_3 = otters_data_clean_v11a$Detail_3
otters_data_clean_v11$Detail_4 = otters_data_clean_v11a$Detail_4

## STEP 6: MISCELLANEOUS DATA CLEANING ##
otters_data_clean_v12 = otters_data_clean_v11 %>%
  # Nest by entry-to-exit sequence
  group_by(entry_id) %>%
  nest() %>%
  # For each entry-to-exit sequence...
  mutate(data = map(data, ~.x %>%
                      # Convert dump outs to exits
                      mutate(Event = ifelse(Event == "Dump In/Out" & lag(Event) == "Turnover", "Exit", Event)) %>%
                      # Convert goals from defensive zone (mainly empty netters) into exits
                      mutate(Event = ifelse(Event == "Goal" & lag(Event) == "Turnover" & Team == lag(Team), "Exit", Event)) %>%
                      # Determine the next and previous event for each event
                      mutate(prev_event = lag(Event), next_event = lead(Event)))) %>%
  # Unnest the data
  unnest() %>% ungroup()


## STEP 7: FORMATTING DATA INTO EASY-TO-SUBSET DATA FRAME FOR MODELLING ##
otters_data_clean_v13 = otters_data_clean_v12 %>%
  # Create a binary variable to identify when the game switches
  mutate(game_switch = ifelse(game_date != lag(game_date), 1, 0)) %>%
  mutate(game_switch = ifelse(game_switch %in% NA, 0, game_switch)) %>%
  # Use this to create a game_id variable
  mutate(game_id = cumsum(game_switch) + 1) %>%
  # Identify the defensive team
  mutate(def_team = ifelse(entry_team == Home_Team, Away_Team, Home_Team)) %>%
  # Convert home team and away team goals to offensive and defensive team goals
  mutate(offensive_team_score = ifelse(entry_team == Home_Team, Home_Team_Goals, Away_Team_Goals),
         defensive_team_score = ifelse(entry_team == Home_Team, Away_Team_Goals, Home_Team_Goals)) %>%
  # Calculate the offensive score difference
  mutate(offensive_score_diff = offensive_team_score - defensive_team_score) %>%
  # Calculate the offensive strength difference (ie PP, ES, PK)
  mutate(offensive_strength_diff = offensive_team_skaters - defensive_team_skaters) %>%
  # Calculate the cumulative seconds into the game
  mutate(c_seconds = (Period - 1) * 1200 + Total_Seconds) %>%
  # Select our desired columns and order them
  select(game_id, episode_id, entry_id, 
         period = Period, c_seconds, seconds = Total_Seconds, 
         event = Event, prev_event, next_event, 
         entry_team, def_team, poss_team = Team, Player, Player_2, 
         same_direction_x, same_direction_y, same_direction_x2, same_direction_y2, 
         detail_1 = Detail_1, detail_2 = Detail_2, detail_3 = Detail_3, detail_4 = Detail_4, 
         last_event_b4_faceoff, time_since_entry, offensive_score_diff, offensive_strength_diff) %>%
  # Remove a couple of straggler observations that aren't associated with an entry
  filter(!(entry_id %in% NA))


## STEP 8: FURTHER PROCESSING BASED ON OUR ASSUMPTIONS ##
otters_data_clean_v14 = otters_data_clean_v13 %>%
  # Remove a couple of odd cases where "Dump In/Out" and "Zone Entry" occur at the exact same time but the dump in is listed first
  # so it counts towards the previous sequence. These coincidentally align with shots as the prev_event
  filter(!(event == "Dump In/Out" & prev_event == "Shot")) %>%
  # Convert mid-sequence dump-ins to passes
  mutate(event = ifelse(event == "Dump In/Out" & prev_event != "Zone Entry" & detail_1 == "Retained", "Pass", event)) %>%
  mutate(event = ifelse(event == "Dump In/Out" & prev_event != "Zone Entry" & detail_1 == "Lost", "Failed Pass", event)) %>%
  mutate(detail_1 = ifelse(event %in% c("Pass", "Failed Pass") & detail_1 %in% c("Retained", "Lost"), "Indirect", detail_1)) %>%
  # Remove recoveries after a possession play
  filter(!(event == "Recovery" & prev_event %in% c("Pass", "Recovery"))) %>%
  # Remove recoveries after non-dumped zone entries
  filter(!(event == "Recovery" & prev_event == "Zone Entry" & lag(detail_1 %in% c("Carried", "Played")))) %>%
  # Combine controlled zone entry then dump in to dumped zone entry
  mutate(detail_1 = ifelse(event == "Zone Entry" & next_event == "Dump In/Out", "Dumped", detail_1),
         same_direction_x2 = ifelse(event == "Zone Entry" & next_event == "Dump In/Out", lead(same_direction_x), same_direction_x2),
         same_direction_y2 = ifelse(event == "Zone Entry" & next_event == "Dump In/Out", lead(same_direction_y), same_direction_y2)) %>%
  # Remove dump in events after zone entries (since we consider the zone entry 'dumped' already)
  filter(!(event == "Dump In/Out" & prev_event == "Zone Entry")) %>%
  # Recalculate prev_event and next_event
  group_by(entry_id) %>%
  nest() %>%
  mutate(data = future_map(data, ~.x %>% mutate(prev_event = lag(event), next_event = lead(event)))) %>%
  unnest() %>%
  # Turn next_event into Whistle for frozen shots
  mutate(next_event = ifelse(event == "Shot" & last_event_b4_faceoff == 1, "Whistle", next_event)) %>%
  # Remove exits (only needed as next_event)
  filter(event != "Exit") %>%
  # Remove overtimes
  filter(period != 4) %>%
  # Nest data by entry-to-exit sequence
  group_by(entry_id) %>%
  nest() %>%
  # Get number of events in each sequence
  mutate(len = future_map(data, ~nrow(.x)) %>% as.numeric()) %>%
  # Remove sequences of length 1
  filter(len > 1) %>%
  select(-len) %>%
  # Unnest, ungroup
  unnest() %>% ungroup() %>%
  # Label strength state as ES/PP/SH
  mutate(strength_state = case_when(
    offensive_strength_diff == 0 ~ "ES",
    offensive_strength_diff > 0 ~ "PP",
    offensive_strength_diff < 0 ~ "SH"
  ))


## STEP 9: ADDITIONAL DATA CLEANING ##
otters_data_clean_v15 = otters_data_clean_v14 %>%
  # Add in a score effect classification
  mutate(score_effect = case_when(
    offensive_score_diff >= 1 & period == 3 ~ "Lead",
    offensive_score_diff >= 2 & period < 3 ~ "Lead",
    offensive_score_diff <= -1 & period == 3 ~ "Deficit",
    offensive_score_diff <= -2 & period < 3 ~ "Deficit",
    TRUE ~ "Close"
  )) %>%
  # Nest data by entry-to-exit sequence
  group_by(entry_id) %>%
  nest() %>%
  # For each entry-to-exit sequence, get the strength state at the entry
  mutate(entry_strength = map(data, ~head(.x, 1) %>% select(strength_state) %>% unlist())) %>%
  # Unnest
  unnest() %>% ungroup() %>%
  # Remove SH plays
  filter(entry_strength != "SH") %>%
  # Identify if the previous event was a pass
  mutate(is_prev_event_pass = ifelse(event %in% c("Shot", "Goal") & prev_event %in% c("Pass", "Failed Pass"), 1, 0)) %>%
  # Determine where the pass location and time was if the last event was a pass
  mutate(last_pass_x = ifelse(is_prev_event_pass == 1, lag(same_direction_x), NA),
         last_pass_y = ifelse(is_prev_event_pass == 1, lag(same_direction_y), NA),
         time_since_last_pass = ifelse(is_prev_event_pass == 1, c_seconds - lag(c_seconds), NA)) %>%
  # Fix what appears to be a data entry error (entry at x = 28)
  mutate(same_direction_x = ifelse(same_direction_x == 28 & event == "Zone Entry", 128, same_direction_x)) %>%
  # Remove shots from behind the net (we assume that cannot happen in our model)
  filter(!(event %in% c("Shot", "Goal") & same_direction_x > 189)) %>%
  # If we converted a dump-in event to a pass (ie it occurred >25 feet into the zone), then we relabel the preceding entry as a carried entry
  mutate(detail_1 = ifelse(detail_1 == "Dumped" & next_event %in% c("Failed Pass", "Pass"), "Carried", detail_1)) %>%
  # Add in score effect
  mutate(score_effect = case_when(
    (period %in% c(1,2)) & (offensive_score_diff >= 2) ~ "Lead",
    (period %in% c(1,2)) & (offensive_score_diff <= -2) ~ "Deficit",
    TRUE ~ "Close"
  ))
```

# Preparing Data for Simulations and Analysis

Here we perform add in a few extra variables and clean up some details to prepare our data for simulations and analyses.

```{r simulations}
## STEP 1: ASSIGN TURNOVERS TO THE PLAYER THAT TURNS THE PUCK OVER ##
otters_data_clean_v16 = otters_data_clean_v15 %>%
  # Turnovers are assigned to the last (known) player who had the puck
  mutate(Player = ifelse(event == "Turnover", lag(Player), Player)) %>%
  # Make sure that this is Player_2 for complete passes
  mutate(Player = ifelse(event == "Turnover" & prev_event == "Pass", lag(Player_2), Player))


## STEP 2: DETERMINE WHICH EVENTS ARE ASSIGNED TO OTTERS ##

# Get a list of all Otters in the data
otters_players = otters_data_clean_v16 %>%
  filter(poss_team == "Erie Otters") %>%
  select(Player) %>%
  unique() %>%
  unlist()

# Create a variable called is_otter to identify them
otters_data_clean_v17 = otters_data_clean_v16 %>%
  mutate(is_otter = ifelse(Player %in% otters_players, 1, 0))


## STEP 3: ADD IN PASS COUNTS ##

# Create an event ID
otters_data_clean_v18 = otters_data_clean_v17 %>% mutate(event_id = 1:nrow(.))

# Get pass counts
pass_data = otters_data_clean_v18 %>%
  select(event_id, entry_id, event, prev_event) %>%
  # Nest by entry-to-exit sequence
  group_by(entry_id) %>%
  nest() %>%
  # For each entry-to-exit sequence...
  mutate(data = map(data, ~.x %>% 
                      # Create a pass sequence ID to identify plays with consecutive passes
                      mutate(num = ifelse(!(event %in% c("Pass", "Failed Pass")) & prev_event %in% c("Pass", "Failed Pass"), 1, 0)) %>%
                      mutate(pass_seq_id = cumsum(num)))) %>%
  # Unnest
  unnest() %>%
  # Nest by each pass subsequence within each entry-to-exit sequence
  group_by(entry_id, pass_seq_id) %>%
  nest() %>%
  # For each pass subsequence...
  mutate(data = map(data, ~.x %>% 
                      # Remove non-passes
                      filter(event %in% c("Pass", "Failed Pass")) %>%
                      # Create a variable that records the cumulative number of passes in a subsequence
                      mutate(num2 = 1) %>%
                      mutate(pass_count = cumsum(num2)))) %>%
  # Unnest
  unnest() %>% ungroup() %>%
  # If there are more than 5 passes, label it as 5
  mutate(pass_count = ifelse(pass_count >= 4, 4, pass_count))

# Merge the pass counts into the main data frame
otters_data_clean_v19 = otters_data_clean_v18 %>%
  merge(pass_data, on = c("event_id"), all.x = TRUE) %>%
  mutate(pass_count = ifelse(pass_count %in% NA, 0, pass_count)) %>%
  arrange(event_id)


## STEP 4: ADD IN SHOT COUNTS AND TIME SINCE LAST SHOT ##
otters_data_clean_v20 = otters_data_clean_v19 %>%
  # Nest by entry-to-exit sequence
  group_by(entry_id) %>%
  nest() %>%
  # For each entry-to-exit sequence...
  mutate(data = map(data, ~.x %>%
                      # ... Determine the total number of shots through the play so far at each event
                      mutate(dummy = ifelse(event == "Shot", 1, 0)) %>%
                      mutate(shot_count = cumsum(dummy))
                      )) %>%
  # Unnest and ungroup
  unnest() %>%
  ungroup() %>%
  # Nest by entry-to-shot and shot-to-shot subsequences
  group_by(entry_id, shot_count) %>%
  nest() %>%
  # For each subsequence...
  mutate(data = map(data, ~.x %>%
                      # ... Determine how long it has been since the last shot
                      mutate(s_sec = .x$seconds[1]) %>%
                      mutate(time_since_last_shot = seconds - s_sec))) %>%
  # Unnest and ungroup
  unnest() %>%
  ungroup() %>%
  # Time since last shot is NA if a shot has not occurred yet
  mutate(time_since_last_shot = ifelse(shot_count == 0, NA, time_since_last_shot)) %>%
  # Time since last shot is NA when the first shot occurs
  mutate(time_since_last_shot = ifelse(shot_count == 1 & event == "Shot", NA, time_since_last_shot)) %>%
  # Time since last shot is the time since the previous shot for the 2nd+ shot in a sequence
  mutate(time_since_last_shot = ifelse(shot_count > 1 & event == "Shot", seconds - lag(s_sec), time_since_last_shot))


## STEP 5: CREATE TIME-DECAYING VARIABLES FOR TIME SINCE LAST SHOT/PASS/ENTRY ##
otters_data_clean_v21 = otters_data_clean_v20 %>%
  # Create a decaying time function for time since the entry
  mutate(entry_time_decay = case_when(
    # 0 seconds since entry --> 1
    time_since_entry == 0 ~ 1,
    # 1-9 seconds since entry --> 1 - log10(seconds)
    time_since_entry < 10 ~ 1 - log10(time_since_entry),
    # 10+ seconds since entry --> 0
    time_since_entry >= 10 ~ 0,
    TRUE ~ 0
  )) %>%
  # Create a decaying time function for time since the last shot
  mutate(shot_time_decay = case_when(
    # 0 seconds since last shot --> 1
    time_since_last_shot == 0 ~ 1,
    # 1-9 seconds since last shot --> 1 - log10(seconds)
    time_since_last_shot < 10 ~ 1 - log10(time_since_last_shot),
    # 10+ seconds since last shot --> 0
    time_since_last_shot >= 10 ~ 0,
    TRUE ~ 0
  )) %>%
  # Create a decaying time function for time since the last pass
  mutate(pass_time_decay = case_when(
    # 0 seconds since last pass --> 1
    time_since_last_pass == 0 ~ 1,
    # 1-9 seconds since last pass --> 1 - log10(seconds)
    time_since_last_pass < 10 ~ 1 - log10(time_since_last_pass),
    # 10+ seconds since last pass --> 0
    time_since_last_pass >= 10 ~ 0,
    TRUE ~ 0
  ))
  

## STEP 5: FINAL TOUCH UPS ##
otters_data_clean_v22 = otters_data_clean_v21 %>%
  # Remove useless variables
  select(-s_sec, -num, -num2, -dummy, -pass_seq_id) %>%
  # Add in a zone entry indicator
  mutate(before_entry_indicator = ifelse(event == "Zone Entry", 1, 0)) %>%
  # Organize column order
  select(game_id, episode_id, entry_id, event_id, event, prev_event, next_event, everything(), shot_count)
```


Save the final data used in our models as a csv.

```{r save3}
write.csv(otters_data_clean_v22, here::here("Data/otters_data_for_modelling.csv"), row.names = FALSE)
```
