---
output: pdf_document
---

\centering
\Huge

**INLA SPDE Model - xG Sample Model Based off Events of Shots**

\raggedright

\rule{\textwidth}{1.5pt}
\small

```{r setup, include = FALSE}
library(knitr) # For formatting the document
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

Summary: This file provides a template for building our spatiotemporal logistic regression model with INLA to determine the expected goals of a given shot that was NOT preceded by a pass.


# Load data and packages

Load necessary packages

```{r load packages}
library(tidyverse) # For various data analysis functions

library(here) # To call or save code/data in certain folders

library(INLA) # To fit model with INLA

source(here::here("Code/2. Functions/Plotting-Functions.R")) # For ggField2 function
```

Load data set

```{r load data}


samples2 = readRDS(here::here("Data/all_samples_so_far2.Rda")) %>%
  filter(prev_event %in% c("Shot", "Play", "Incomplete Play", "Zone Entry", "Turnover", "Recovery"))


```


# Data Preparation

Create additional variables including a binary variable to identify goals (our response variable), a count of the total shots in a sequence, as well as decaying time functions for time since last shot and time since last entry.

```{r data prep}

otters_data = read_csv(here::here("Data/otters_data_for_modelling.csv")) %>%
  filter(entry_strength == "ES") %>%
  mutate(time_since_entry.x = time_since_entry) %>%
  filter(event != "Zone Entry")

samples2 = samples2 %>%
  mutate(xG_0 = ifelse(Expected_Goals == 0, 1, 0)) %>%
  filter(xG_0 == 0) %>%
  bind_rows(otters_data) %>%
  mutate(last_minute = ifelse(seconds > 1140, 1, 0)) %>%
  mutate(time_last_minute = ifelse(last_minute == 1, abs(seconds - 1200), 0)) %>%
  mutate(is_lead = ifelse(score_effect == "Lead", 1, 0)) %>%
  mutate(is_deficit = ifelse(score_effect == "Deficit", 1, 0)) %>%
  mutate(shot_time_decay = case_when(
    # 0 seconds since last shot --> 1
    time_since_last_shot.x == 0 ~ 1,
    # 1-9 seconds since last shot --> 1 - log10(seconds)
    time_since_last_shot.x < 10 ~ 1 - log10(time_since_last_shot.x),
    # 10+ seconds since last shot --> 0
    time_since_last_shot.x >= 10 ~ 0,
    TRUE ~ 0
  )) %>%
  mutate(is_pass = ifelse(prev_event %in% c("Play", "Incomplete Play"), 1, 0)) %>%
  mutate(is_shot = ifelse(prev_event %in% "Shot", 1, 0)) %>%
  mutate(is_turnover = ifelse(prev_event %in% "Turnover", 1, 0)) %>%
  mutate(is_recovery = ifelse(prev_event %in% "Recovery", 1, 0)) %>%
  mutate(is_behind_net = ifelse(same_direction_x > 189, 1, 0))
```


# Create Meshes

Build the spatial shot mesh.

```{r shot mesh}
oz_boundary = data.frame(
  x = c(125, 200, 200, 125, 125),
  y = c(0, 0, 85, 85, 0)
)


model_domain = as.matrix(oz_boundary)
model_coords2 = as.matrix(samples2 %>% select(same_direction_x, same_direction_y))

# Create the mesh
Mesh = inla.mesh.2d(model_coords2, loc.domain = model_domain, max.edge = c(18, 12), offset = c(1,2), cutoff = 8)

# Plot the mesh along with the data points
plot(Mesh)
points(model_coords2, col = "blue", pch = 4)

# Save the mesh
#saveRDS(Mesh, here::here(paste("Models/Meshes/Shot_to_Goal_no_pass", strength, "Shot_Mesh.Rda", sep = "_")))
```

Build the temporal mesh for time since the zone entry occurred.

```{r time mesh}
# Set the knots
knots = c(0, 2, 5, 10, max(samples2$time_since_entry.x))

# Create 1D mesh
time_Mesh <- inla.mesh.1d(loc = knots)

# Save the mesh
#saveRDS(time_Mesh, here::here(paste("Models/Meshes/Shot_to_Goal_no_pass", strength, "Time_Mesh.Rda", sep = "_")))
```



# Model Specifications

Set a prior and create a weight matrix and index for the spatial effect in our model.

```{r spatial shot}
# Create a weight matrix based on our spatial and temporal meshes
model_A_matrix2 = inla.spde.make.A(Mesh, loc = model_coords2, n.group = length(knots), group = samples2$time_since_entry.x, group.mesh = time_Mesh)

# Set the PC Matern prior
model_spde2 = inla.spde2.pcmatern(mesh = Mesh, prior.range = c(20, 0.8), prior.sigma = c(0.5, 0.5))

# Create the weight index
model_w_index2 = inla.spde.make.index("w", n.spde = model_spde2$n.spde, n.group = length(knots))
```

Create our model matrix.

```{r model matrix}
X2 = samples2 %>%
  # Select necessary columns to be used in model
  select(is_deficit, is_lead, shot_time_decay, shot_count, time_last_minute, last_minute, is_behind_net,
         is_pass, is_shot, is_turnover, is_recovery, pass_count)
```

Create a stack of all information to include in the model.

```{r stack}
# Initialize the stack
model_stack2 = inla.stack(
  
  # Add our response variable
  data = list(y = samples2$Expected_Goals), 
  
  # Add a list of the weights on the effects
  A = list(1, 1, model_A_matrix2),
  
  # Add spatial effects
  effects = list(
    
    # Intercept, repeat 1 n times
    Intercept = rep(1, nrow(samples2)),
    
    # Add our model matrix
    X = X2,
    
    # Add spatial effect
    model_w_index2
    
  ))
```

State our model formula.

```{r formula}
model_formula2 = as.formula(paste0("y ~ -1 + Intercept ", 
                                  " + shot_time_decay:is_shot + shot_count + is_lead + is_deficit + time_last_minute + last_minute",
                                  " + pass_count:is_pass + is_behind_net",
                                  " + is_pass + is_shot + is_recovery + is_turnover",
                                  " + f(w, model = model_spde2, group = w.group, control.group = list(model = 'ar1'))"))
```


# Model Results

Run the model

```{r model}
model2 = inla(model_formula2, 
             
             family = "gamma",
             
             data = inla.stack.data(model_stack2),
             
             control.compute = list(dic = TRUE, config = TRUE),
             
             control.predictor = list(A = inla.stack.A(model_stack2)),
             
             num.threads = 8
)
```

Evaluate the model

```{r viz}
# Create the spatial map
model_map = gg_OZ(model2, Mesh, Groups = 5,
                  GroupLabels = c("0 Seconds Since Entry", "2 Seconds Since Entry", 
                                  "5 Seconds Since Entry", "10 Seconds Since Entry",
                                  ">10 Seconds Since Entry")) +
  scale_fill_viridis_d()

# Print a model summary
summary(model2)
```

Save our results

```{r save}
inz_gamma_model = model2
saveRDS(inz_gamma_model, file = here::here("Models/Value/In_Zone_Gamma_Model.Rda"))
```

```{r}
fitted_values_gamma = inz_gamma_model$summary.fitted.values[1:nrow(samples2),]

fitted_df_gamma = (cbind(samples2, fitted_values_gamma) %>%
  filter(is.na(Expected_Goals)) %>%
  mutate(mean = exp(mean)) %>%
  select(event_id, gamma = mean))[1:19546,]


fitted_df = merge(fitted_df_binom, fitted_df_gamma, by = "event_id", all.x = TRUE)

fitted_df = fitted_df %>%
  mutate(xV = (1-mean) * gamma)

```



