---
output: pdf_document
---

\centering
\Huge

**INLA SPDE Model - Controlled Entry Transition Model 1D**

\raggedright

\rule{\textwidth}{1.5pt}
\small

```{r setup, include = FALSE}
library(knitr) # For formatting the document
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

Summary: This file provides a template for building our spatiotemporal poisson regression model with INLA used in the transition matrix of our Markov Decision Process.

In particular, we are fitting the following transition models:

- Controlled Entry to Pass (to run, set to = c("Play", "Incomplete Play") on line 58)

- Controlled Entry to Shot (to run, set to = c("Shot", "Goal") on line 58)

- Controlled Entry to Turnover (to run, set to = c("Turnover") on line 58)


# Load data and packages

Load necessary packages

```{r load packages}
library(tidyverse) # For various data analysis functions

library(here) # To call or save code/data in certain folders

library(INLA) # To fit model with INLA

source(here::here("Code/2. Functions/Plotting-Functions.R")) # For ggField2 function
```

Load data set

```{r load data}
samples = readRDS(here::here("Data/all_samples_so_far2.Rda")) %>%
  filter(!is.na(event_id)) %>%
  filter(event == "Zone Entry")
```

# Data Preparation

Prepare our model matrix.

```{r data prep}
otters_data = read_csv(here::here("Data/otters_data_for_modelling.csv")) %>%
  filter(entry_strength == "ES") %>%
  mutate(time_since_entry.x = time_since_entry) %>%
  filter(event == "Zone Entry")

samples = samples %>%
  mutate(xG_0 = ifelse(Expected_Goals == 0, 1, 0)) %>%
  bind_rows(otters_data) %>%
  mutate(is_near_boards = ifelse(same_direction_y <= 10 | same_direction_y >= 75, 1, 0)) %>%
  mutate(is_lead = ifelse(score_effect == "Lead", 1, 0)) %>%
  mutate(is_deficit = ifelse(score_effect == "Deficit", 1, 0)) %>%
  mutate(last_minute = ifelse(seconds > 1140, 1, 0)) %>%
  mutate(time_last_minute = ifelse(last_minute == 1, abs(seconds - 1200), 0))

model_coords = samples$same_direction_y
```


# Building the INLA SPDE Mesh

Create a mesh

```{r mesh}
Mesh = inla.mesh.1d(0:85)
```

# Set Model Specifications

Set a prior and create a weight matrix and index for our model.

```{r spatial effect}
# Create an A matrix (sparse matrix representing each triangle)
model_A_matrix = inla.mesh.1d.A(Mesh, loc = model_coords)

# Set the PC Matern prior
model_spde = inla.spde2.pcmatern(mesh = Mesh, prior.range = c(50, 0.6), prior.sigma = c(0.5, 0.5))

# Create w index
model_w_index = inla.spde.make.index("i", n.spde = model_spde$n.spde)
```


```{r}
X = samples %>% 
  select(is_lead, is_deficit, time_last_minute, last_minute, is_near_boards)
```

Create a stack of all information to include in the model.

```{r stack}
# Initialize the stack
model_stack = inla.stack(
  
  # Add our response variable
  data = list(y = samples$xG_0), 
  
  # Add a list of the weights on the effects
  A = list(1, 1, model_A_matrix),
  
  # Add the effects
  effects = list(
    
    # Intercept, repeat 1 n times
    Intercept = rep(1, nrow(samples)),
    
    # Create a variable called is_near_boards to help us deal with discontinuity at the boards
    X = X,
    
    # Our 1D spatial effect
    model_w_index
    
  ))
```

Set the model formula

```{r formula}
model_formula = as.formula(paste0("y ~ -1 + Intercept ", 
                                  " + is_near_boards + is_lead + is_deficit + last_minute + time_last_minute",
                                  " + f(i, model = model_spde)"))
```

# Model Results

Run our model

```{r model}
model = inla(model_formula,
             
             family = "binomial",
             
             data = inla.stack.data(model_stack),
             
             control.compute = list(dic = TRUE, config = TRUE),
             
             control.predictor = list(A = inla.stack.A(model_stack)),
             
             num.threads = 8
)
```

Evaluate our model

```{r viz}
# Extract fitted values
fitval = model$summary.fitted.values[1:nrow(samples),] %>%
  mutate(x = samples$same_direction_y)

# Plot entry means
entry_1d = ggplot(fitval) +
  geom_point(aes(x = x, y = mean)) +
  geom_smooth(aes(x = x, y = mean), se = FALSE, colour = "green", size = 2)



# Print a model summary
summary(model)
```




```{r}
ze_binom_model = model

#saveRDS(ze_binom_model, here::here("Models/Value/Zone_Entry_Binomial_Model.Rda"))
```

```{r}
fitted_values_binom_ze = ze_binom_model$summary.fitted.values[1:8759,]

fitted_df_binom_ze = cbind(samples, fitted_values_binom_ze) %>%
  filter(is.na(xG_0))
```


