---
output: pdf_document
---

\centering
\Huge

**INLA SPDE Model - Traffic Probability Model on Shots with Preceding Pass with AR(1) Process for Time Since Entry**

\raggedright

\rule{\textwidth}{1.5pt}
\small

```{r setup, include = FALSE}
library(knitr) # For formatting the document
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

Summary: This file provides a template for building our spatiotemporal logistic regression model with INLA used in the simulations of our model. Here we are predicting the probability of traffic on a given shot that is preceded by a pass.


# Load data and packages

Load necessary packages

```{r load packages}
library(tidyverse) # For various data analysis functions

library(here) # To call or save code/data in certain folders

library(INLA) # To fit model with INLA

source(here::here("Code/2. Functions/Plotting-Functions.R")) # For custom plotting functions
```

Load data set

```{r load data}
otters_data = read_csv(here::here("Data/otters_data_for_modelling.csv"))
```

# Specify Strength State

Specify which strength state we wish to run the model in

```{r parameters}
strength = "ES" # "PP" or "ES"
```


# Data Preparation

Filter down to the data we wish to build the model with (shots that are preceded by a pass).

```{r data prep}
# Trim to our desired events (shots)
shot_data = otters_data %>%
  # Filter for shots
  filter(event %in% c("Shot", "Goal")) %>%
  # Filter for specified strength state
  filter(entry_strength == strength) %>%
  # Filter for shots that are preceded by a pass
  filter(is_prev_event_pass == 1) %>%
  # Convert logical variables to 0-1 binaries
  mutate(detail_3 = as.numeric(detail_3), 
         detail_4 = as.numeric(detail_4))
```


# Create Meshes

Build the spatial shot mesh.

```{r shot mesh}
# Create a coordinate matrix
locations = shot_data %>% 
  select(same_direction_x, same_direction_y) %>%
  mutate_all(as.numeric) %>%
  as.matrix()

# Set the boundary for the mesh
xvals = locations[,1]
yvals = locations[,2]

xrange = c(min(xvals), 189)
yrange = c(min(yvals), 85)

model_domain <- matrix(c(xrange[1], yrange[1],
                         xrange[2], yrange[1],
                         xrange[2], yrange[2],
                         xrange[1], yrange[2],
                         xrange[1], yrange[1]),
                       ncol = 2L, byrow = TRUE)


# Create the mesh
Mesh = inla.mesh.2d(locations, loc.domain = model_domain, max.edge = c(20, 15), offset = c(1, 2), cutoff = 8)


# Plot the mesh and data points
plot(Mesh)
points(locations, col = "blue", pch = 4)

# Save the mesh
saveRDS(Mesh, here::here(paste("Models/Meshes/Traffic_with_pass", strength, "Shot_Mesh.Rda", sep = "_")))
```

Build the spatial pass mesh.

```{r pass mesh}
# Create a coordinate matrix
pass_loc = shot_data %>% 
  select(last_pass_x, last_pass_y) %>%
  mutate_all(as.numeric) %>%
  as.matrix()

# Set the boundary for the mesh
pass_x = pass_loc[,1]
pass_y = pass_loc[,2]

pass_xrange = c(min(pass_x), 200)
pass_yrange = c(min(pass_y), 85)

pass_domain <- matrix(c(pass_xrange[1], pass_yrange[1],
                        pass_xrange[2], pass_yrange[1],
                        pass_xrange[2], pass_yrange[2],
                        pass_xrange[1], pass_yrange[2],
                        pass_xrange[1], pass_yrange[1]),
                      ncol = 2L, byrow = TRUE)

# Create the mesh
pass_Mesh = inla.mesh.2d(pass_loc, loc.domain = pass_domain, max.edge = c(20, 15), offset = c(1, 2), cutoff = 8)

# Plot the mesh
plot(pass_Mesh)
points(pass_loc, col = "red", pch = 4)

saveRDS(pass_Mesh, here::here(paste("Models/Meshes/Traffic_with_pass", strength, "Pass_Mesh.Rda", sep = "_")))
```

Build the temporal mesh.

```{r time mesh}
# Set the knots
knots = c(0, 2, 5, 10, max(shot_data$time_since_entry))

# Create 1D meshes
time_Mesh <- inla.mesh.1d(loc = knots)

# Save the meshes
saveRDS(time_Mesh, here::here(paste("Models/Meshes/Traffic_with_pass", strength, "Time_Mesh.Rda", sep = "_")))
```

# Model Specifications

Set a prior and create a weight matrix and index for the shot spatial effect in our model.

```{r spatial shot}
# Create a spatial weight matrix for shot locations
model_A_matrix = inla.spde.make.A(Mesh, loc = locations, n.group = length(knots), group = shot_data$time_since_entry, group.mesh = time_Mesh)

# Set the PC Matern prior
model_spde = inla.spde2.pcmatern(mesh = Mesh, prior.range = c(25, 0.6), prior.sigma = c(0.5, 0.5))

# Create shot spatial weight index
model_w_index = inla.spde.make.index("w", n.spde = model_spde$n.spde, n.group = length(knots))
```

Set a prior and create a weight matrix and index for the pass spatial effect in our model.

```{r spatial pass}
# Create a spatial weight matrix for pass locations
pass_A_matrix = inla.spde.make.A(pass_Mesh, loc = pass_loc, n.group = length(knots), group = shot_data$time_since_entry, group.mesh = time_Mesh)

# Set the PC Matern prior
pass_spde = inla.spde2.pcmatern(mesh = pass_Mesh, prior.range = c(40, 0.6), prior.sigma = c(0.5, 0.5))

# Create pass spatial weight index
pass_w_index = inla.spde.make.index("g", n.spde = pass_spde$n.spde, n.group = length(knots))
```

Create a stack of all information to include in the model.

```{r stack}
# Initialize the stack
model_stack = inla.stack(
  
  # Add our response variable
  data = list(y = shot_data$detail_3), 
  
  # Add a list of the weights on the effects
  A = list(1, pass_A_matrix, model_A_matrix),
  
  # Add spatial effects
  effects = list(
    
    # Intercept, repeat 1 n times
    Intercept = rep(1, nrow(shot_data)),
    
    # Add pass spatial effect
    pass_w_index,
    
    # Add shot spatial effect
    model_w_index
    
  ))
```

State our model formula.

```{r formula}
model_formula = as.formula(paste0("y ~ -1 + Intercept ", 
                                  " + f(w, model = model_spde, group = w.group, control.group = list(model = 'ar1'))",
                                  " + f(g, model = pass_spde, group = g.group, control.group = list(model = 'ar1'))"))
```


# Model Results

Run the model

```{r model}
model = inla(model_formula,
             
             family = "binomial",
             
             data = inla.stack.data(model_stack),
             
             control.compute = list(dic = TRUE, config = TRUE),
             
             control.predictor = list(A = inla.stack.A(model_stack)),
             
             num.threads = 8
)
```

Evaluate the model

```{r viz}
# Create the spatial maps
model_map_shot = gg_OZ(model, Mesh, Mat_Name = "w", Groups = 5, family = "binomial",
                  GroupLabels = c("0 Seconds Since Entry", "2 Seconds Since Entry", 
                                  "5 Seconds Since Entry", "10 Seconds Since Entry",
                                  ">10 Seconds Since Entry")) +
  labs(title = paste("Traffic Probability on Shot without Pass ", strength, " AR(1) INLA Spatial Map", sep = ""))

model_map_pass = gg_OZ(model, Mesh, Mat_Name = "w", Groups = 5, family = "binomial",
                  GroupLabels = c("0 Seconds Since Entry", "2 Seconds Since Entry", 
                                  "5 Seconds Since Entry", "10 Seconds Since Entry",
                                  ">10 Seconds Since Entry")) +
  labs(title = paste("Pass Before Shot with Traffic ", strength, " AR(1) INLA Spatial Probability Map", sep = ""))

# Save the spatial maps
ggsave(filename = paste("Traffic Probability on Shot with Pass ", strength, " AR(1) INLA Spatial Map.png", sep = ""), 
       plot = model_map_shot, device = "png", path = here::here("Figures/Glue Models/"))

ggsave(filename = paste("Pass Before Shot with Traffic ", strength, " AR(1) INLA Spatial Probability Map.png", sep = ""), 
       plot = model_map_pass, device = "png", path = here::here("Figures/Glue Models/"))

# Print a model summary
summary(model)
```

Save our results

```{r save}
saveRDS(model, here::here(paste("Models/Glue Models/Traffic_with_pass", strength, "model.Rda", sep = "_")))
```



