---
output: pdf_document
---

\centering
\Huge

**INLA SPDE Model - XY-Location Prediction Model**

\raggedright

\rule{\textwidth}{1.5pt}
\small

```{r setup, include = FALSE}
library(knitr) # For formatting the document
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

Summary: This file provides a template for building our spatiotemporal regression model with INLA used in the simulations of our model. Here we are predicting the next xy-coordinates following an event.


# Load data and packages

Load necessary packages

```{r load packages}
library(tidyverse) # For various data analysis functions

library(here) # To call or save code/data in certain folders

library(INLA) # To fit model with INLA

source(here::here("Code/2. Functions/Plotting-Functions.R")) # For ggField2 function
```

Load data set

```{r load data}
otters_data = read_csv(here::here("Data/otters_data_for_modelling.csv"))
```

# Specify Strength State

Specify which strength state we wish to run the model in

```{r parameters}
strength = "ES" # "PP" or "ES"
```


# Data Preparation

Filter down to the data we wish to build the model with (shots that are preceded by a pass).

```{r data prep}
# Add necessary variables
otters_data = otters_data %>%
  # Filter for strength state
  filter(entry_strength == strength) %>%
  # Create binary variables for previous events
  mutate(prev_play = ifelse(prev_event %in% c("Play", "Incomplete Play"), 1, 0)) %>%
  mutate(prev_shot = ifelse(prev_event == "Shot", 1, 0)) %>%
  mutate(prev_recovery = ifelse(prev_event == "Recovery", 1, 0)) %>%
  mutate(prev_turnover = ifelse(prev_event == "Turnover", 1, 0)) %>%
  mutate(prev_zone_entry = ifelse(prev_event == "Zone Entry", 1, 0)) %>%
  # Create binary for direct passes
  mutate(direct_pass = case_when(
    is.na(detail_1) ~ 0,
    detail_1 == "Direct" ~ 1,
    TRUE ~ 0
  )) %>%
  # Nest by entry-to-exit sequence
  group_by(entry_id) %>%
  nest() %>%
  # For each entry-to-exit sequence...
  mutate(data = map(data, ~.x %>% 
                      mutate(prev_x = lag(same_direction_x)) %>%
                      mutate(prev_y = lag(same_direction_y)))) %>%
  # Unnest
  unnest() %>%
  # Determine if the previous event came from the left side
  mutate(prev_left_side = ifelse(prev_y > 42.5, 1, 0)) %>%
  # Determine if the previous event came from behind the net
  mutate(prev_behind_net = ifelse(prev_x >= 189, 1, 0)) %>%
  # Get z-scored x-coordinate
  mutate(x_normalized = scale(same_direction_x)) %>%
  # Create interaction term between x_normalized and right side of ice
  mutate(x_normalized_int_right = x_normalized * (1 - prev_left_side)) %>%
  # Filter for events with previous events (no zone entries)
  filter(!is.na(prev_event)) %>%
  # Relabel pass events to direct and indirect
  mutate(event2 = case_when(
    event %in% c("Play", "Incomplete Play") & detail_1 == "Direct" ~ "Direct",
    event %in% c("Play", "Incomplete Play") & detail_1 == "Indirect" ~ "Indirect",
    TRUE ~ event
  )) %>%
  # Assign each event type a numeric value
  mutate(type_index = plyr::mapvalues(event2, from = c("Turnover","Recovery","Shot", "Direct", "Indirect"), to 1:5) %>% as.numeric())
```


# Create Meshes

Build the spatial event mesh.

```{r shot mesh}
# Create the boundary for OZ mesh
oz_boundary = data.frame(
  x = c(125, 200, 200, 125, 125),
  y = c(0, 0, 85, 85, 0)
)

model_domain = as.matrix(oz_boundary)


# Get model coordinates and matrix
model_coords = as.matrix(otters_data %>% select(prev_x, prev_y))
model_matrix = otters_data %>% select(prev_x, prev_y, type_index)


# Build INLA SPDE mesh
Mesh = inla.mesh.2d(model_coords, max.edge = c(20,20), offset = 40, cutoff = 20, boundary = model_domain) # Set for computation ease rather than because it is a great mesh, fairly coarse


# Plot the mesh and data points
plot(Mesh)
points(locations, col = "blue", pch = 4)


# Save the mesh
saveRDS(Mesh, here::here(paste("Models/Meshes/Mesh_x_y", strength, "prediction.Rda", sep = "_")))
```


# Model Specifications

Set a prior and create a weight matrix and index for the spatial effect in our model.

```{r spatial shot}
# Create a spatial weight matrix for shot locations
model_A_matrix = inla.spde.make.A(Mesh, loc = model_coords, n.group = 5, group = otters_data$type_index)

# Set the PC Matern prior
model_spde = inla.spde2.pcmatern(mesh = Mesh, prior.range = c(20, 0.2), prior.sigma = c(0.5, 0.5))

# Create shot spatial weight index
model_w_index = inla.spde.make.index("w", n.spde = model_spde$n.spde, n.group = 5)
```


Create a matrix of all covariates used in the model

```{r model matrix}
X = otters_data %>%
  select(direct_pass, prev_behind_net, prev_play, prev_shot, prev_recovery, prev_turnover, time_since_entry,
         x_normalized, x_normalized_int_right)
```


# X-Location Model

Create a stack of all information to include in the model.

```{r stack}
# Initialize the stack
model_stack_x = inla.stack(
  
  # Add our response variable
  data = list(y = otters_data$same_direction_x), 
  
  # Add a list of the weights on the effects
  A = list(1, 1, model_A_matrix),
  
  # Add spatial effects
  effects = list(
    
    # Intercept, repeat 1 n times
    Intercept = rep(1, nrow(shot_data)),
    
    # Add our model matrix
    X = X,
    
    # Add shot spatial effect
    model_w_index
    
  ))
```

State our model formula.

```{r formula}
model_formula_x = as.formula(paste0("y ~ -1 + Intercept ", 
                                    " + direct_pass + prev_behind_net + time_since_entry",
                                    " + prev_play + prev_recovery + prev_shot + prev_zone_entry",
                                    " + f(w, model = model_spde, group = w.group, control.group = list(model = 'iid'))"))
```


Run the model

```{r model}
model_x = inla(model_formula_x,
             
             family = "gaussian",
             
             data = inla.stack.data(model_stack_x),
             
             control.compute = list(dic = TRUE, config = TRUE),
             
             control.predictor = list(A = inla.stack.A(model_stack_x)),
             
             num.threads = 16
)
```


Evaluate the model

```{r viz}
# Create the spatial maps
model_map_x = gg_OZ(model_x, Mesh, Mat_Name = "w", Groups = 5, 
                    GroupLabels = c("Turnover", "Recovery", "Shot", "Direct Pass", "Indirect Pass")) +
  labs(title = paste("X-Location ", strength, " IID INLA Spatial Map", sep = "")) +
  scale_fill_viridis_d()

# Save the spatial maps
ggsave(filename = paste("X-Location ", strength, " IID INLA Spatial Map.png", sep = ""), 
       plot = model_map_x, device = "png", path = here::here("Figures/Miscellaneous/"))

# Print a model summary
summary(model)
```

Save our results

```{r save}
saveRDS(model_x, here::here(paste("Models/Miscellaneous/X_Location", strength, "model.Rda", sep = "_")))
```


# Y-Location Model (Dependent on X)

Create a stack of all information to include in the model.

```{r stack}
# Initialize the stack
model_stack_y = inla.stack(
  
  # Add our response variable
  data = list(y = otters_data$same_direction_y), 
  
  # Add a list of the weights on the effects
  A = list(1, 1, model_A_matrix),
  
  # Add spatial effects
  effects = list(
    
    # Intercept, repeat 1 n times
    Intercept = rep(1, nrow(shot_data)),
    
    # Add our model matrix
    X = X,
    
    # Add shot spatial effect
    model_w_index
    
  ))
```

State our model formula.

```{r formula}
model_formula_y = as.formula(paste0("y ~ -1 + Intercept ", 
                                    " + prev_behind_net + time_since_entry + x_normalized + x_normalized_int_right",
                                    " + prev_play + prev_recovery + prev_shot + prev_zone_entry",
                                    " + f(w, model = model_spde, group = w.group, control.group = list(model = 'iid'))"))
```


Run the model

```{r model}
model_y = inla(model_formula_y,
             
             family = "gaussian",
             
             data = inla.stack.data(model_stack_y),
             
             control.compute = list(dic = TRUE, config = TRUE),
             
             control.predictor = list(A = inla.stack.A(model_stack_y)),
             
             num.threads = 16
)
```


Evaluate the model

```{r viz}
# Create the spatial maps
model_map_y = gg_OZ(model_y, Mesh, Mat_Name = "w", Groups = 5,
                    GroupLabels = c("Turnover", "Recovery", "Shot", "Direct Pass", "Indirect Pass")) +
  labs(title = paste("Y-Location ", strength, " IID INLA Spatial Map", sep = "")) +
  scale_fill_viridis_d()

# Save the spatial maps
ggsave(filename = paste("Y-Location ", strength, " IID INLA Spatial Map.png", sep = ""), 
       plot = model_map_y, device = "png", path = here::here("Figures/Miscellaneous/"))

# Print a model summary
summary(model)
```

Save our results

```{r save}
saveRDS(model_y, here::here(paste("Models/Miscellaneous/Y_Location", strength, "model.Rda", sep = "_")))
```

