---
output: pdf_document
---

\centering
\Huge

**INLA SPDE Model - Controlled Entry Transition Model 1D**

\raggedright

\rule{\textwidth}{1.5pt}
\small

```{r setup, include = FALSE}
library(knitr) # For formatting the document
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

Summary: This file provides a template for building our spatiotemporal poisson regression model with INLA used in the transition matrix of our Markov Decision Process.

In particular, we are fitting the following transition models:

- Controlled Entry to Pass (to run, set to = c("Play", "Incomplete Play") on line 58)

- Controlled Entry to Shot (to run, set to = c("Shot", "Goal") on line 58)

- Controlled Entry to Turnover (to run, set to = c("Turnover") on line 58)


# Load data and packages

Load necessary packages

```{r load packages}
library(tidyverse) # For various data analysis functions

library(here) # To call or save code/data in certain folders

library(INLA) # To fit model with INLA

source(here::here("Code/2. Functions/Plotting-Functions.R")) # For ggField2 function
```

Load data set

```{r load data}
otters_data = read_csv(here::here("Data/otters_data_for_modelling.csv"))
```


# Specify Event Transition

Specify the event transition we wish to model along with the strength state.

```{r parameters}
to = c("Play")

strength = "ES"  # "PP" or "ES"
```

# Data Preparation

Prepare our model matrix.

```{r data prep}
# Narrow down to plays that meet our specifications
otters_data_filtered = otters_data %>%
  # Select our desired start and end plays
  filter(event == "Zone Entry" & detail_1 != "Dumped" & next_event %in% to) %>%
  # Select our desired strength state
  filter(strength_state == strength)


# Create a table containing all available y-coordinate values along with their event counts in each group
otters_data_table = otters_data_filtered %>%
  # Select xy-coordinates and time since entry (grouped)
  select(same_direction_y) %>%
  # Convert factors to numeric values
  mutate_all(as.character) %>%
  mutate_all(as.numeric) %>%
  # Convert data frame to a table of counts for each possible combination of xy-coordinates and entry_time_group
  table() %>%
  as.data.frame()

# Convert the column names to "y_coord" and "y" (our response)
colnames(otters_data_table) = c("y_coord", "y")


# Some values are still missing, create a table with all possible y-coordinates we can enter
model_matrix = data.frame(y_coord = rep(0:85)) %>%
  # Merge in the counts from otters_data_table
  merge(otters_data_table, by = c("y_coord"), all = TRUE) %>%
  # Replace NAs with 0s
  mutate(y = ifelse(y %in% NA, 0, y))


# Create a matrix of all y-coordinates to be used in the creation of our mesh
model_coords = model_matrix %>%
  # Select just xy-coordinates
  select(y_coord) %>%
  mutate_all(as.character) %>%
  mutate_all(as.numeric) %>%
  # Convert to a vector
  unlist()
```


# Building the INLA SPDE Mesh

Create a mesh

```{r mesh}
Mesh = inla.mesh.1d(model_coords)
```

# Set Model Specifications

Set a prior and create a weight matrix and index for our model.

```{r spatial effect}
# Create an A matrix (sparse matrix representing each triangle)
model_A_matrix = inla.mesh.1d.A(Mesh, loc = model_coords)

# Set the PC Matern prior
model_spde = inla.spde2.pcmatern(mesh = Mesh, prior.range = c(50, 0.6), prior.sigma = c(0.5, 0.5))

# Create w index
model_w_index = inla.spde.make.index("i", n.spde = model_spde$n.spde)
```

Create a stack of all information to include in the model.

```{r stack}
# Initialize the stack
model_stack = inla.stack(
  
  # Add our response variable
  data = list(y = model_matrix$y), 
  
  # Add a list of the weights on the effects
  A = list(1, 1, model_A_matrix),
  
  # Add the effects
  effects = list(
    
    # Intercept, repeat 1 n times
    Intercept = rep(1, nrow(model_matrix)),
    
    # Create a variable called is_near_boards to help us deal with discontinuity at the boards
    X = data.frame(is_near_boards = c(rep(1,10), rep(0,66), rep(1,10))),
    
    # Our 1D spatial effect
    model_w_index
    
  ))
```

Set the model formula

```{r formula}
model_formula = as.formula(paste0("y ~ -1 + Intercept ", 
                                  " + is_near_boards",
                                  " + f(i, model = model_spde)"))
```

# Model Results

Run our model

```{r model}
model = inla(model_formula,
             
             family = "poisson",
             
             data = inla.stack.data(model_stack),
             
             control.compute = list(dic = TRUE, config = TRUE),
             
             control.predictor = list(A = inla.stack.A(model_stack)),
             
             num.threads = 8
)
```

Evaluate our model

```{r viz}
# Extract fitted values
fitval = model$summary.fitted.values[1:86,] %>%
  mutate(x = 0:85)

# Plot entry means
entry_1d = ggplot(fitval) +
  geom_line(aes(x = x, y = mean)) +
  labs(title = paste("Controlled Entry to", to[1], strength, "1D INLA Mesh", sep = " "),
       x = "Location Along Blue Line", y = "Mean Count")

# Save the spatial map
ggsave(filename = paste("Controlled Entry to", to[1], strength, "1D INLA Mesh.png", sep = " "), plot = entry_1d, 
       device = "png", path = here::here("Figures/Entry/"))

# Print a model summary
summary(model)
```

Save our results

```{r save}
# Save the model
saveRDS(model, file = here::here("Models/Entry", paste("Controlled_Entry_to", to[1], strength, "model.Rda", sep = "_")))
```







