---
output: pdf_document
---

\centering
\Huge

**INLA SPDE Model - Recovery, Shot and Turnover Transition Models with AR(1) Process for Time Since Entry**

\raggedright

\rule{\textwidth}{1.5pt}
\small

```{r setup, include = FALSE}
library(knitr) # For formatting the document
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

Summary: This file provides a template for building our spatiotemporal poisson regression model with INLA used in the transition matrix of our Markov Decision Process.

In particular, we are fitting the following transition models:

- Recovery to Shot (to run, set from = c("Recovery"); to = c("Shot", "Goal") on lines 68 and 70, respectively)

- Recovery to Pass (to run, set from = c("Recovery"); to = c("Play", "Incomplete Play") on lines 68 and 70, respectively)

- Recovery to Turnover (to run, set from = c("Recovery"); to = c("Turnover") on lines 68 and 70, respectively)

- Shot to Recovery (to run, set from = c("Shot"); to = c("Recovery") on lines 68 and 70, respectively)

- Shot to Turnover (to run, set from = c("Shot"); to = c("Turnover") on lines 68 and 70, respectively)

- Shot to Whistle (to run, set from = c("Shot"); to = c("Whistle") on lines 68 and 70, respectively)

- Turnover to Exit (to run, set from = c("Turnover"); to = c("Exit") on lines 68 and 70, respectively)

- Turnover to Recovery (to run, set from = c("Turnover"); to = c("Recovery") on lines 68 and 70, respectively)


# Load data and packages

Load necessary packages

```{r load packages}
library(tidyverse) # For various data analysis functions

library(here) # To call or save code/data in certain folders

library(INLA) # To fit model with INLA

source(here::here("Code/2. Functions/Plotting-Functions.R")) # For plotting function
```

Load data set

```{r load data}
otters_data = read_csv(here::here("Data/otters_data_for_modelling.csv"))
```


# Specify Event Transition

Specify the event transition we wish to model along with the strength state.

```{r parameters}
from = c("Recovery")

to = c("Play", "Incomplete Play")

strength = "ES"
```


# Data Preparation

Extract point counts at each xy-coordinate and each time bin.

```{r data prep}
# Select our desired start and end plays
otters_data_filtered = otters_data %>%
  # Filter for our desired transition and strength state
  filter(event %in% from & next_event %in% to) %>%
  filter(strength_state == strength) %>%
  # Bin time since entry that will be used as the temporal effect in our model
  mutate(entry_time_group = case_when(
    # 0-2 seconds
    time_since_entry <= 2 ~ 1,
    # 3-5 seconds
    time_since_entry <= 5 ~ 2,
    # 6-10 seconds
    time_since_entry <= 10 ~ 3,
    # 10+ seconds
    TRUE ~ 4
  ))

# Create a table containing all available x and y values along with their event counts in each group
otters_data_table = otters_data_filtered %>%
  # Select xy-coordinates and time since entry (grouped)
  select(same_direction_x, same_direction_y, entry_time_group) %>%
  # Convert factors to numeric values
  mutate_all(as.character) %>%
  mutate_all(as.numeric) %>%
  # Convert data frame to a table of counts for each possible combination of xy-coordinates and entry_time_group
  table() %>%
  as.data.frame() %>%
  # Rename variables
  dplyr::rename(x_coord = same_direction_x, y_coord = same_direction_y, y = Freq)
```

Create data frames that capture the offensive zone boundary and all points in the OZ as well as a model matrix that includes all possible combinations of x coordinates, y coordinates and entry time bin, along with the frequency of event occurrences in each bin.

```{r model matrix}
if (from == "Shot") {
  
  # Set a boundary around the OZ
  oz_boundary = data.frame(
  x = c(125, 189, 189, 125, 125),
  y = c(0, 0, 85, 85, 0)
  )
  
  # Select all values in the OZ
  oz_full = data.frame(x_coord = rep(seq(125,189), 86), y_coord = rep(0:85, each = 65))
  
  
  # Replicate the oz_full data frame 4 times
  model_matrix = do.call("rbind", replicate(4, oz_full, simplify = FALSE)) %>%
    # For each replicate, add in the corresponding entry time bin number (1-4)
    mutate(entry_time_group = rep(1:4, each = 65*86)) %>%
    # Merge in the frequencies
    merge(otters_data_table, by = c("x_coord", "y_coord", "entry_time_group"), all = TRUE) %>%
    # Replace NAs with 0s
    mutate(y = ifelse(y %in% NA, 0, y))
    
} 

if (from != "Shot") {
  
  # Set a boundary around the OZ
  oz_boundary = data.frame(
  x = c(125, 200, 200, 125, 125),
  y = c(0, 0, 85, 85, 0)
  )
  
  # Select all values in the OZ
  oz_full = data.frame(x_coord = rep(seq(125,200), 86), y_coord = rep(0:85, each = 76))
  
  
  # Replicate the oz_full data frame 4 times
  model_matrix = do.call("rbind", replicate(4, oz_full, simplify = FALSE)) %>%
    # For each replicant, add in the corresponding entry time bin number (1-4)
    mutate(entry_time_group = rep(1:4, each = 76*86)) %>%
    # Merge in the frequencies
    merge(otters_data_table, by = c("x_coord", "y_coord", "entry_time_group"), all = TRUE) %>%
    # Replace NAs with 0s
    mutate(y = ifelse(y %in% NA, 0, y))
}
```


# Building the INLA SPDE Mesh

Create a mesh

```{r mesh}
# Set the domain and coordinates to be used in our mesh
model_domain = as.matrix(oz_boundary)
model_coords = as.matrix(model_matrix[1:2])

# Create the mesh
Mesh = inla.mesh.2d(model_coords, loc.domain = model_domain, max.edge = c(20, 15), offset = c(1,2), cutoff = 8)

# Plot the mesh along with the data points
plot(Mesh)
points(model_coords, col = "blue", pch = 4)

if (from == "Shot") {
  saveRDS(Mesh, here::here("Models/Meshes/Shot_Transition_Mesh.Rda"))
}

if (from != "Shot") {
  saveRDS(Mesh, here::here("Models/Meshes/Recov_Turnover_Transition_Mesh.Rda"))
}
```

# Set Model Specifications

Set a prior and create a weight matrix and index for the spatial effect in our model.

```{r spatial effect}
# Create a spatial weight matrix
model_A_matrix = inla.spde.make.A(Mesh, loc = model_coords,
                                  group = model_matrix$entry_time_group,
                                  n.group = 4)

# Set the PC Matern prior
model_spde = inla.spde2.pcmatern(mesh = Mesh, prior.range = c(10, 0.8), prior.sigma = c(0.5, 0.5))

# Create spatial weight index
model_w_index = inla.spde.make.index("w", n.spde = model_spde$n.spde, n.group = 4)
```

Create a stack of all information to include in the model.

```{r stack}
# Initialize the stack
model_stack = inla.stack(
  
  # Add our response variable
  data = list(y = model_matrix$y), 
  
  # Add a list of the weights on the effects
  A = list(1, model_A_matrix),
  
  # Add spatial effects
  effects = list(
    
    # Intercept, repeat 1 n times
    Intercept = rep(1, nrow(model_matrix)),
    
    # Add spatial effect
    model_w_index
    
  ))
```

Set the model formula

```{r formula}
model_formula = as.formula(paste0("y ~ -1 + Intercept + f(w, model = model_spde, group = w.group, control.group = list(model = 'ar1'))"))
```

# Model Results

Run our model

```{r model}
model = inla(model_formula,
             
             family = "poisson",
             
             data = inla.stack.data(model_stack),
             
             control.compute = list(dic = TRUE, config = TRUE),
             
             control.predictor = list(A = inla.stack.A(model_stack)),
             
             num.threads = 8
)
```

Evaluate our model

```{r viz}
# Visualize the spatial effects with a spatial heatmap
model_map = gg_OZ(model, Mesh, Groups = 4, family = "poisson",
                  GroupLabels = c("0-2 Seconds After Entry", "3-5 Seconds After Entry", 
                                  "6-10 Seconds After Entry", "11+ Seconds After Entry")) +
  labs(title = paste(from, "to", to[1], strength, "Entry Time AR(1) INLA Spatial Map", sep = " "))

# Save the spatial map
ggsave(filename = paste(from, "to", to[1], strength, "AR(1) INLA Spatial Map.png", sep = " "), plot = model_map, device = "png", 
       path = here::here(paste("Figures/", from, "/", sep = "")))

# Print a model summary
summary(model)
```

Save our results as an Rda file.

```{r save}
# Save the model
saveRDS(model, file = here::here(paste("Models/", from, sep = ""), paste(from, "to", to[1], strength, "model.Rda", sep = "_")))
```










