---
output: pdf_document
---

\centering
\Huge

**INLA SPDE Model - Shot to Goal Probability Model without Pre-Shot Pass**

\raggedright

\rule{\textwidth}{1.5pt}
\small

```{r setup, include = FALSE}
library(knitr) # For formatting the document
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

Summary: This file provides a template for building our spatiotemporal logistic regression model with INLA to determine the expected goals of a given shot that was NOT preceded by a pass.


# Load data and packages

Load necessary packages

```{r load packages}
library(tidyverse) # For various data analysis functions

library(here) # To call or save code/data in certain folders

library(INLA) # To fit model with INLA

source(here::here("Code/2. Functions/Plotting-Functions.R")) # For ggField2 function
```

Load data set

```{r load data}
otters_data = read_csv(here::here("Data/otters_data_for_modelling.csv"))
```


# Specify Strength State

Specify which strength state we wish to run the model in.

```{r parameters}
strength = "ES"
```


# Data Preparation

Create additional variables including a binary variable to identify goals (our response variable), a count of the total shots in a sequence, as well as decaying time functions for time since last shot and time since last entry.

```{r data prep}
# Trim data and create additional variables
shot_data = otters_data %>%
  # Filter for shots
  filter(event %in% c("Shot", "Goal")) %>%
  # Nest by entry-to-exit sequence
  group_by(entry_id) %>%
  nest() %>%
  # For each entry-to-exit sequence...
  mutate(data = map(data, ~.x %>% 
                      # Create a shot count variable to tally the cumulative number of shots in a play
                      mutate(num = 1) %>%
                      mutate(shot_count = cumsum(num)) %>%
                      # Determine the time since the last shot
                      mutate(last_shot_time = c_seconds - lag(c_seconds)) %>%
                      # Mark first shots
                      mutate(is_first_shot = ifelse(shot_count == 1, 1, 0)))) %>%
  # Unnest
  unnest() %>% ungroup() %>%
  # Create a decaying time function for time since the entry
  mutate(entry_time_decay = case_when(
    # 0 seconds since entry --> 1
    time_since_entry == 0 ~ 1,
    # 1-9 seconds since entry --> 1 - log10(seconds)
    time_since_entry < 10 ~ 1 - log10(time_since_entry),
    # 10+ seconds since entry --> 0
    time_since_entry >= 10 ~ 0,
    TRUE ~ 0
  )) %>%
  # Create a decaying time function for time since the last shot
  mutate(shot_time_decay = case_when(
    # 0 seconds since last shot --> 1
    last_shot_time == 0 ~ 1,
    # 1-9 seconds since last shot --> 1 - log10(seconds)
    last_shot_time < 10 ~ 1 - log10(last_shot_time),
    # 10+ seconds since last shot --> 0
    last_shot_time >= 10 ~ 0,
    TRUE ~ 0
  )) %>%
  # Create a binary response variable
  mutate(is_goal = ifelse(event == "Goal", 1, 0)) %>%
  # Convert logical variables to 0-1 binaries
  mutate(detail_3 = as.numeric(detail_3), 
         detail_4 = as.numeric(detail_4))

# Filter for data matching this transition...
shot_data = shot_data %>%
  # ... no pass occurs directly before shot
  filter(is_prev_event_pass == 0) %>%
  # ... only our specified strength state
  filter(entry_strength == strength)
```


# Create Meshes

Build the spatial shot mesh.

```{r shot mesh}
# Create a coordinate matrix
locations = shot_data %>% 
  select(same_direction_x, same_direction_y) %>%
  mutate_all(as.character) %>%
  mutate_all(as.numeric) %>%
  as.matrix()

# Set the boundary for the mesh
xvals = locations[,1]
yvals = locations[,2]

xrange = c(125, 189)
yrange = c(0, 85)

model_domain <- matrix(c(xrange[1], yrange[1],
                         xrange[2], yrange[1],
                         xrange[2], yrange[2],
                         xrange[1], yrange[2],
                         xrange[1], yrange[1]),
                       ncol = 2L, byrow = TRUE)


# Create the mesh
Mesh = inla.mesh.2d(locations, loc.domain = model_domain, max.edge = c(20, 15), offset = c(1, 2), cutoff = 8)


# Plot the mesh and data points
plot(Mesh)
points(locations, col = "blue", pch = 4)

# Save the mesh
saveRDS(Mesh, here::here(paste("Models/Meshes/Shot_to_Goal_no_pass", strength, "Shot_Mesh.Rda", sep = "_")))
```

Build the temporal mesh for time since the zone entry occurred.

```{r time mesh}
# Set the knots
knots = c(0, 2, 5, 10, max(shot_data$time_since_entry))

# Create 1D mesh
time_Mesh <- inla.mesh.1d(loc = knots)

# Save the mesh
saveRDS(time_Mesh, here::here(paste("Models/Meshes/Shot_to_Goal_no_pass", strength, "Time_Mesh.Rda", sep = "_")))
```



# Model Specifications

Set a prior and create a weight matrix and index for the spatial effect in our model.

```{r spatial shot}
# Create a weight matrix based on our spatial and temporal meshes
model_A_matrix = inla.spde.make.A(Mesh, loc = locations, n.group = length(knots), group = shot_data$time_since_entry, group.mesh = time_Mesh)

# Set the PC Matern prior
model_spde = inla.spde2.pcmatern(mesh = Mesh, prior.range = c(20, 0.8), prior.sigma = c(0.5, 0.5))

# Create the weight index
model_w_index = inla.spde.make.index("w", n.spde = model_spde$n.spde, n.group = length(knots))
```

Create our model matrix.

```{r model matrix}
X = shot_data %>%
  # Select necessary columns to be used in model
  select(offensive_score_diff, offensive_strength_diff, period, shot_time_decay, shot_count,
         detail_3, is_first_shot, is_prev_event_pass, time_since_entry)
```

Create a stack of all information to include in the model.

```{r stack}
# Initialize the stack
model_stack = inla.stack(
  
  # Add our response variable
  data = list(y = shot_data$is_goal), 
  
  # Add a list of the weights on the effects
  A = list(1, 1, model_A_matrix),
  
  # Add spatial effects
  effects = list(
    
    # Intercept, repeat 1 n times
    Intercept = rep(1, nrow(shot_data)),
    
    # Add our model matrix
    X = X,
    
    # Add spatial effect
    model_w_index
    
  ))
```

State our model formula.

```{r formula}
model_formula = as.formula(paste0("y ~ -1 + Intercept ", 
                                  " + shot_time_decay + shot_count + detail_3",
                                  " + f(w, model = model_spde, group = w.group, control.group = list(model = 'ar1'))"))
```


# Model Results

Run the model

```{r model}
model = inla(model_formula, 
             
             family = "binomial",
             
             data = inla.stack.data(model_stack),
             
             control.compute = list(dic = TRUE, config = TRUE),
             
             control.predictor = list(A = inla.stack.A(model_stack)),
             
             num.threads = 8
)
```

Evaluate the model

```{r viz}
# Create the spatial map
model_map_shot = gg_OZ(model, Mesh, Groups = 5, family = "binomial",
                  GroupLabels = c("0 Seconds Since Entry", "2 Seconds Since Entry", 
                                  "5 Seconds Since Entry", "10 Seconds Since Entry",
                                  ">10 Seconds Since Entry")) +
  labs(title = paste("Shot to Goal without Pre-Shot Movement ", strength, " AR(1) INLA Spatial Map", sep = ""))

# Save the spatial maps
ggsave(filename = paste("Shot to Goal without Pre-Shot Pass ", strength, " AR(1) INLA Spatial Map.png", sep = ""), 
       plot = model_map_shot, device = "png", path = here::here("Figures/Shot/"))

# Print a model summary
summary(model)
```

Save our results

```{r save}
# Save the model
saveRDS(model, file = here::here("Models/Shot/", paste("Shot_to_Goal", strength, "no_pass_model.Rda", sep = "_")))
```

Extract and save fitted expected goals values from the model.

```{r}
# Get xG fitted values
xG_fitted = model$summary.fitted.values[1:nrow(shot_data),]

# Combine with the observed shot data
shot_data = cbind(shot_data, xG_fitted)

# Save as a csv
write.csv(shot_data, here::here("Data/no_pass_xG.csv"), row.names = FALSE)
```


