---
output: pdf_document
---

\centering
\Huge

**INLA SPDE Model - Dump-In Entry Transition Models with AR(1) Process for Score Effects**

\raggedright

\rule{\textwidth}{1.5pt}
\small

```{r setup, include = FALSE}
library(knitr) # For formatting the document
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

Summary: This file provides a template for building our spatiotemporal poisson regression model with INLA used in the transition matrix of our Markov Decision Process.

In particular, we are fitting the following transition models:

- Dumped Entry to Recovery (to run, set to = c("Recovery") on line 58)

- Dumped Entry to Turnover (to run, set to = c("Turnover") on line 58)


# Load data and packages

Load necessary packages

```{r load packages}
library(tidyverse) # For various data analysis functions

library(here) # To call or save code/data in certain folders

library(INLA) # To fit model with INLA

source(here::here("Code/0. Functions/Plotting-Functions.R")) # For ggField2 function
```

Load data set

```{r load data}
otters_data = read_csv(here::here("Data/otters_data_for_modelling.csv"))
```


# Specify Event Transition

Specify the event transition we wish to model along with the strength state.

```{r parameters}
to = c("Turnover")

strength = "ES" # "PP" or "ES"
```


# Data Preparation

Extract point counts at each xy-coordinate and each score effect bin.

```{r data prep}
# Select our desired start and end plays
otters_data_filtered = otters_data %>%
  # Filter for our desired transition and strength state
  filter(event == "Zone Entry" & detail_1 == "Dumped" & next_event %in% to) %>%
  filter(entry_strength == strength) %>%
  # Replace entry coordinates with dump-in coordinates
  mutate(same_direction_x = ifelse(!(same_direction_x2 %in% NA), same_direction_x2, same_direction_x),
         same_direction_y = ifelse(!(same_direction_y2 %in% NA), same_direction_y2, same_direction_y)) %>%
  # Remove dump-ins from the defensive zone
  filter(same_direction_x >= 75) %>%
  # Remove any dump-ins that occur > 25 feet into the zone (that isn't really an entry anymore)
  filter(same_direction_x < 150) %>%
  # Convert score effects (score close) to a numeric value (1 = Deficit, 2 = Close, 3 = Lead)
  mutate(score_effect2 = factor(score_effect, levels = c("Deficit", "Close", "Lead")) %>% as.numeric())

# Create a table containing all available x and y values along with their event counts in each group
otters_data_table = otters_data_filtered %>%
  # Select xy-coordinates and score effect
  select(same_direction_x, same_direction_y, score_effect2) %>%
  # Convert factors to numeric values
  mutate_all(as.character) %>%
  mutate_all(as.numeric) %>%
  # Convert data frame to a table of counts for each possible combination of xy-coordinates and entry_time_group
  table() %>%
  as.data.frame() %>%
  # Rename variables
  dplyr::rename(x_coord = same_direction_x, y_coord = same_direction_y, y = Freq)
```

Create our model matrix as well as the boundary in which the spatial effects have to lie.

```{r model matrix}
# Set a boundary around where dump-ins can occur
entry_boundary = data.frame(
  x = c(75, 150, 150, 75, 75),
  y = c(0, 0, 85, 85, 0)
)

# Select all of poissble dump-in entries
entry_full = data.frame(x_coord = rep(seq(75,150), 86), y_coord = rep(0:85, each = 76))


# Replicate the oz_full data frame 3 times
model_matrix = do.call("rbind", replicate(3, entry_full, simplify = FALSE)) %>%
  # For each replicate, add in the corresponding score state
  mutate(score_effect2 = rep(1:3, each = 76*86)) %>%
  # Merge in the frequencies
  merge(otters_data_table, by = c("x_coord", "y_coord", "score_effect2"), all = TRUE) %>%
  # Replace NAs with 0s
  mutate(y = ifelse(y %in% NA, 0, y))
```


# Building the INLA SPDE Mesh

Create a mesh

```{r mesh}
# Set the domain and coordinates to be used in our mesh
model_domain = as.matrix(entry_boundary)
model_coords = as.matrix(model_matrix[1:2])

# Create the mesh
Mesh = inla.mesh.2d(model_coords, loc.domain = model_domain, max.edge = c(20, 15), offset = c(1,2), cutoff = 8)

# Plot the mesh along with the data points
plot(Mesh)
points(model_coords, col = "blue", pch = 4)

# Save the mesh for future use
saveRDS(Mesh, here::here("Models/Meshes/Dump_In_Entry_Mesh.Rda"))
```

# Set Model Specifications

Set a prior and create a weight matrix and index for our model.

```{r spatial effect}
# Create a weight matrix
model_A_matrix = inla.spde.make.A(Mesh, loc = model_coords,
                                  group = model_matrix$score_effect2,
                                  n.group = 3)

# Set the PC Matern prior
model_spde = inla.spde2.pcmatern(mesh = Mesh, prior.range = c(3, 0.8), prior.sigma = c(0.5, 0.5))

# Create weight index
model_w_index = inla.spde.make.index("w", n.spde = model_spde$n.spde, n.group = 3)
```

Create a stack of all information to include in the model.

```{r stack}
# Initialize the stack
model_stack = inla.stack(
  
  # Add our response variable
  data = list(y = model_matrix$y), 
  
  # Add a list of the weights on the effects
  A = list(1, model_A_matrix),
  
  # Add spatial effects
  effects = list(
    
    # Intercept, repeat 1 n times
    Intercept = rep(1, nrow(model_matrix)),
    
    # Add spatial effect
    model_w_index
    
  ))
```

Set the model formula

```{r formula}
model_formula = as.formula(paste0("y ~ -1 + Intercept ", 
                                  " + f(w, model = model_spde, group = w.group, control.group = list(model = 'ar1'))"))
```

# Model Results

Run our model

```{r model}
model = inla(model_formula,
             
             family = "poisson",
             
             data = inla.stack.data(model_stack),
             
             control.compute = list(dic = TRUE, waic = TRUE, config = TRUE),
             
             control.predictor = list(A = inla.stack.A(model_stack)),
             
             num.threads = 8
)
```

Evaluate our model

```{r viz}
# Visualize the spatial effects with a spatial heatmap
model_map = gg_NZ(model, Mesh, Groups = 3, family = "poisson",
                  GroupLabels = c("Down 2+ in 1st/2nd, Down 1+ in 3rd", "Within 1 in 1st/2nd, Tied in 3rd", 
                                  "Up 2+ in 1st/2nd, Up 1+ in 3rd")) +
  labs(title = paste("Dump-In to", to[1], strength, "Score Effect AR(1) INLA Spatial Map", sep = " ")) +
  xlim(c(75, 150))

# Save the spatial map
ggsave(filename = paste("Dump-In to", to[1], strength, "Score Effect AR(1) INLA Spatial Map.png", sep = " "), plot = model_map, 
       device = "png", path = here::here("Figures/Entry/"))

# Print a model summary
summary(model)
```

Save our results

```{r save}
# Save the model
saveRDS(model, file = here::here("Models/Entry", paste("Dump_to", to[1], strength, "model.Rda", sep = "_")))
```







