---
output: pdf_document
---

\centering
\Huge

**INLA SPDE Model - Shot to Goal Probability Model with Pre-Shot Pass**

\raggedright

\rule{\textwidth}{1.5pt}
\small

```{r setup, include = FALSE}
library(knitr) # For formatting the document
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

Summary: This file provides a template for building our spatiotemporal logistic regression model with INLA to determine the expected goals of a given shot that was preceded by a pass.


# Load data and packages

Load necessary packages

```{r load packages}
library(tidyverse) # For various data analysis functions

library(here) # To call or save code/data in certain folders

library(INLA) # To fit model with INLA

library(raster, exclude = "select") # For distance calculations

library(LearnGeom) # For angle calculations

source(here::here("Code/0. Functions/Plotting-Functions.R")) # For ggField2 function
```

Load data set

```{r load data}
otters_data = read_csv(here::here("Data/otters_data_for_modelling.csv"))
```

# Specify Strength State

Specify which strength state we wish to run the model in

```{r parameters}
strength = "ES"
```


# Data Preparation

Create additional variables including a binary variable to identify goals (our response variable), a count of the total shots in a sequence, as well as decaying time functions for time since last shot and time since last entry. Additionally, calculate pass distance and angle change that will be used as covariates in our model.

```{r data prep}
# Trim data and create additional variables
shot_data = otters_data %>%
  # Filter for shots
  filter(event %in% c("Shot", "Goal")) %>%
  # Create a binary response variable
  mutate(is_goal = ifelse(event == "Goal", 1, 0)) %>%
  # Convert logical variables to 0-1 binaries
  mutate(detail_3 = as.numeric(detail_3), 
         detail_4 = as.numeric(detail_4)) %>%
  # Add variable for passes that come from below goal line
  mutate(from_behind_net = ifelse(last_pass_x > 189, 1, 0)) %>%
  # Make sure all shots have a pass that occurs right before it
  filter(is_prev_event_pass == 1) %>%
  # Filter for only our specified strength state
  filter(entry_strength == strength)

# Add in extra covariates for pass specifications
shot_data$angle = NA
shot_data$pass_angle = NA
shot_data$angle_change = NA
shot_data$pass_distance = NA

for (i in 1:nrow(shot_data)) {
  if (shot_data$is_prev_event_pass[i] == 1) {
    
    shot_data$pass_distance[i] = pointDistance(c(shot_data$same_direction_x[i], shot_data$same_direction_y[i]), 
                                               c(shot_data$last_pass_x[i], shot_data$last_pass_y[i]),
                                               lonlat = FALSE)
    
    shot_data$angle[i] = Angle(A = c(shot_data$same_direction_x[i], shot_data$same_direction_y[i]), 
                               B = c(189, 42.5), C = c(100, 42.5))
    
    shot_data$pass_angle[i] = Angle(A = c(shot_data$last_pass_x[i], shot_data$last_pass_y[i]), 
                                    B = c(189, 42.5), C = c(100, 42.5))
    
    shot_data$angle[i] = ifelse(shot_data$same_direction_y[i] < 42.5, -shot_data$angle[i], shot_data$angle[i])
    shot_data$pass_angle[i] = ifelse(shot_data$same_direction_y[i] < 42.5, -shot_data$pass_angle[i], shot_data$pass_angle[i])
    
    shot_data$angle_change[i] = abs(shot_data$angle[i] - shot_data$pass_angle[i])
    
  }
}

```



# Create Meshes

Build the spatial shot mesh.

```{r shot mesh}
# Create a coordinate matrix
locations = shot_data %>% 
  select(same_direction_x, same_direction_y) %>%
  mutate_all(as.numeric) %>%
  as.matrix()

# Set the boundary for the mesh
xvals = locations[,1]
yvals = locations[,2]

xrange = c(min(xvals), 189)
yrange = c(min(yvals), 85)

model_domain <- matrix(c(xrange[1], yrange[1],
                         xrange[2], yrange[1],
                         xrange[2], yrange[2],
                         xrange[1], yrange[2],
                         xrange[1], yrange[1]),
                       ncol = 2L, byrow = TRUE)


# Create the mesh
Mesh = inla.mesh.2d(locations, loc.domain = model_domain, max.edge = c(20, 15), offset = c(1, 2), cutoff = 8)


# Plot the mesh and data points
plot(Mesh)
points(locations, col = "blue", pch = 4)

saveRDS(Mesh, here::here(paste("Models/Meshes/Shot_to_Goal_with_pass", strength, "Shot_Mesh.Rda", sep = "_")))
```

Build the spatial pass mesh.

```{r pass mesh}
# Create a coordinate matrix
pass_loc = shot_data %>% 
  select(last_pass_x, last_pass_y) %>%
  mutate_all(as.numeric) %>%
  as.matrix()

# Set the boundary for the mesh
pass_x = pass_loc[,1]
pass_y = pass_loc[,2]

pass_xrange = c(min(pass_x), 200)
pass_yrange = c(min(pass_y), 85)

pass_domain <- matrix(c(pass_xrange[1], pass_yrange[1],
                        pass_xrange[2], pass_yrange[1],
                        pass_xrange[2], pass_yrange[2],
                        pass_xrange[1], pass_yrange[2],
                        pass_xrange[1], pass_yrange[1]),
                      ncol = 2L, byrow = TRUE)

# Create the mesh
pass_Mesh = inla.mesh.2d(pass_loc, loc.domain = pass_domain, max.edge = c(20, 15), offset = c(1, 2), cutoff = 8)

# Plot the mesh
plot(pass_Mesh)
points(pass_loc, col = "red", pch = 4)

saveRDS(pass_Mesh, here::here(paste("Models/Meshes/Shot_to_Goal_with_pass", strength, "Pass_Mesh.Rda", sep = "_")))
```

Build the temporal mesh for time since the zone entry occurred.

```{r time mesh}
# Set the knots
knots = c(0, 2, 5, 10, max(shot_data$time_since_entry))

# Create 1D mesh
time_Mesh <- inla.mesh.1d(loc = knots)

saveRDS(time_Mesh, here::here(paste("Models/Meshes/Shot_to_Goal_with_pass", strength, "Time_Mesh.Rda", sep = "_")))
```



# Model Specifications

Set a prior and create a weight matrix and index for the shot spatial effect in our model.

```{r shot spatial}
# Create a weight matrix
model_A_matrix = inla.spde.make.A(Mesh, loc = locations, n.group = length(knots), group = shot_data$time_since_entry, group.mesh = time_Mesh)

# Set the PC Matern prior
model_spde = inla.spde2.pcmatern(mesh = Mesh, prior.range = c(10, 0.8), prior.sigma = c(0.5, 0.5))

# Create weight index
model_w_index = inla.spde.make.index("w", n.spde = model_spde$n.spde, n.group = length(knots))
```

Set a prior and create a weight matrix and index for the pass spatial effect in our model.

```{r pass spatial}
# Create a weight matrix
pass_A_matrix = inla.spde.make.A(pass_Mesh, loc = pass_loc, n.group = length(knots), group = shot_data$time_since_entry, group.mesh = time_Mesh)

# Set the PC Matern prior
pass_spde = inla.spde2.pcmatern(mesh = pass_Mesh, prior.range = c(40, 0.7), prior.sigma = c(0.5, 0.5))

# Create weight index
pass_w_index = inla.spde.make.index("g", n.spde = pass_spde$n.spde, n.group = length(knots))
```

Create our model matrix.

```{r model matrix}
X = shot_data %>%
  # Select necessary columns
  select(pass_time_decay, shot_time_decay, shot_count, detail_3, detail_4, 
         pass_distance, angle_change, from_behind_net)
```

Create a stack of all information to include in the model.

```{r stack}
# Initialize the stack
model_stack = inla.stack(
  
  # Add our response variable
  data = list(y = shot_data$is_goal), 
  
  # Add a list of the weights on the effects
  A = list(1, 1, pass_A_matrix, model_A_matrix),
  
  # Add spatial effects
  effects = list(
    
    # Intercept, repeat 1 n times
    Intercept = rep(1, nrow(shot_data)),
    
    # Add our model matrix
    X = X,
    
    # Add pass spatial effect
    pass_w_index,
    
    # Add shot spatial effect
    model_w_index
    
  ))
```

State our model formula.

```{r formula}
model_formula = as.formula(paste0("y ~ -1 + Intercept ", 
                                  " + pass_time_decay + pass_distance + angle_change + from_behind_net",
                                  " + shot_time_decay",
                                  " + detail_3 + detail_4",
                                  " + f(w, model = model_spde, group = w.group, control.group = list(model = 'ar1'))",
                                  " + f(g, model = pass_spde, group = g.group, control.group = list(model = 'ar1'))"))
```


# Model Results

Run the model

```{r model}
model = inla(model_formula, 
             
             family = "binomial",
             
             data = inla.stack.data(model_stack),
             
             control.compute = list(dic = TRUE, waic = TRUE, config = TRUE),
             
             control.predictor = list(A = inla.stack.A(model_stack)),
             
             num.threads = 8
)
```

Evaluate our model

```{r viz}
# Create the spatial maps
model_map_shot = gg_OZ(model, Mesh, Mat_Name = "w", Groups = 5, family = "binomial",
                  GroupLabels = c("0 Seconds Since Entry", "2 Seconds Since Entry", 
                                  "5 Seconds Since Entry", "10 Seconds Since Entry",
                                  ">10 Seconds Since Entry")) +
  labs(title = paste("Shot to Goal with Pre-Shot Movement ", strength, " AR(1) INLA Spatial Map", sep = ""))

model_map_pass = gg_OZ(model, pass_Mesh, Mat_Name = "g", Groups = 5, family = "binomial",
                  GroupLabels = c("0 Seconds Since Entry", "2 Seconds Since Entry", 
                                  "5 Seconds Since Entry", "10 Seconds Since Entry",
                                  ">10 Seconds Since Entry")) +
  labs(title = paste("Pass Coordinates Before Shot Event ", strength, " AR(1) INLA Spatial Map", sep = ""))

# Save the spatial maps
ggsave(filename = paste("Shot to Goal with Pre-Shot Pass ", strength, " AR(1) INLA Spatial Map.png", sep = ""), 
       plot = model_map_shot, device = "png", path = here::here("Figures/Shot/"))

ggsave(filename = paste("Pass Coordinates Before Shot Event ", strength, " AR(1) INLA Spatial Map.png", sep = ""), 
       plot = model_map_pass, device = "png", path = here::here("Figures/Shot/"))

# Print a model summary
summary(model)
```

Save our results

```{r save}
# Save the model
saveRDS(model, file = here::here("Models/Shot/", paste("Shot_to_Goal", strength, "with_pass_model.Rda", sep = "_")))
```

Extract and save fitted expected goals values from the model.

```{r}
# Get xG fitted values
xG_fitted = model$summary.fitted.values[1:nrow(shot_data),] %>%
  select(xG = mean)

# Combine with the observed shot data
shot_data = cbind(shot_data, xG_fitted)

# Save as a csv
write.csv(shot_data, here::here("Data/with_pass_xG.csv"), row.names = FALSE)
```


