---
output: pdf_document
---

\centering
\Huge

**INLA SPDE Model - Pass Completion Model**

\raggedright

\rule{\textwidth}{1.5pt}
\small

```{r setup, include = FALSE}
library(knitr) # For formatting the document
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

Summary: This file provides a template for building our spatiotemporal logistic regression model with INLA to determine the probability that a pass is direct (no bounces/deflections/banks).


# Load data and packages

Load necessary packages

```{r load packages}
library(tidyverse) # For various data analysis functions

library(here) # To call or save code/data in certain folders

library(INLA) # To fit model with INLA

source(here::here("Code/0. Functions/Plotting-Functions.R")) # For ggField2 function
```

Load data set

```{r load data}
otters_data = read_csv(here::here("Data/otters_data_for_modelling.csv"))
```

# Specify Strength State

Specify which strength state we wish to run the model in

```{r parameters}
strength = "ES"
```


# Data Preparation

Create additional variables including a binary variable to identify goals (our response variable), a count of the total shots in a sequence, as well as decaying time functions for time since last shot and time since last entry. Additionally, calculate pass distance and angle change that will be used as covariates in our model.

```{r data prep}
# Trim data and create additional variables
pass_data = otters_data %>%
  # Filter for passes
  filter(event %in% c("Pass", "Failed Pass")) %>%
  # Filter for only specified strength state
  filter(entry_strength == strength) %>%
  # Binaries for the previous event
  mutate(prev_pass = ifelse(prev_event %in% c("Pass"), 1, 0)) %>%
  mutate(prev_recovery = ifelse(prev_event %in% c("Recovery"), 1, 0)) %>%
  # Binaries for score state
  mutate(is_lead = ifelse(score_effect == "Lead", 1, 0)) %>%
  mutate(is_deficit = ifelse(score_effect == "Deficit", 1, 0)) %>%
  # Determine if pass is direct or not (1 = direct, 2 = indirect)
  mutate(is_direct = ifelse(detail_1 == "Direct", 1, 0))
```



# Create Meshes

Build the INLA spatial mesh.

```{r shot mesh}
# Create a coordinate matrix
locations = pass_data %>% 
  select(same_direction_x, same_direction_y) %>%
  mutate_all(as.numeric) %>%
  as.matrix()

# Set the boundary for the mesh
xvals = locations[,1]
yvals = locations[,2]

xrange = c(min(xvals), 189)
yrange = c(min(yvals), 85)

model_domain <- matrix(c(xrange[1], yrange[1],
                         xrange[2], yrange[1],
                         xrange[2], yrange[2],
                         xrange[1], yrange[2],
                         xrange[1], yrange[1]),
                       ncol = 2L, byrow = TRUE)

# Create the mesh
Mesh = inla.mesh.2d(locations, loc.domain = model_domain, max.edge = c(20, 15), offset = c(1, 2), cutoff = 8)

# Plot the mesh and data points
plot(Mesh)
points(locations, col = "blue", pch = 4)

saveRDS(Mesh, here::here(paste("Models/Meshes/Direct_Pass", strength, "Mesh.Rda", sep = "_")))
```


# Model Specifications

Set a prior and create a weight matrix and index for the spatial effect in our model.

```{r shot spatial}
# Create a weight matrix
model_A_matrix = inla.spde.make.A(Mesh, loc = locations)

# Set the PC Matern prior
model_spde = inla.spde2.pcmatern(mesh = Mesh, prior.range = c(35, 0.5), prior.sigma = c(0.5, 0.5))

# Create weight index
model_w_index = inla.spde.make.index("w", n.spde = model_spde$n.spde)
```

Create our model matrix.

```{r model matrix}
X = pass_data %>%
  # Select necessary columns
  select(prev_pass, prev_recovery, is_lead, is_deficit)
```

Create a stack of all information to include in the model.

```{r stack}
# Initialize the stack
model_stack = inla.stack(
  
  # Add our response variable
  data = list(y = pass_data$is_direct), 
  
  # Add a list of the weights on the effects
  A = list(1, 1, model_A_matrix),
  
  # Add spatial effects
  effects = list(
    
    # Intercept, repeat 1 n times
    Intercept = rep(1, nrow(pass_data)),
    
    # Add our model matrix
    X = X,
    
    # Add spatial effect
    model_w_index
    
  ))
```

State our model formula.

```{r formula}
model_formula = as.formula(paste0("y ~ -1 + Intercept ", 
                                  " + prev_pass + prev_recovery + is_lead + is_deficit",
                                  " + f(w, model = model_spde, group = w.group, control.group = list(model = 'iid'))"))
```


# Model Results

Run the model

```{r model}
model = inla(model_formula, 
             
             family = "binomial",
             
             data = inla.stack.data(model_stack),
             
             control.compute = list(dic = TRUE, waic = TRUE, config = TRUE),
             
             control.predictor = list(A = inla.stack.A(model_stack)),
             
             num.threads = 8
)
```

Evaluate our model

```{r viz}
# Create the spatial maps
model_map = gg_OZ(model, Mesh, Mat_Name = "w", family = "binomial") +
  labs(title = paste("Direct Pass ", strength, " INLA Spatial Map", sep = ""))

# Save the spatial maps
ggsave(filename = paste("Direct Pass ", strength, " INLA Spatial Map.png", sep = ""), 
       plot = model_map, device = "png", path = here::here("Figures/Pass/"))

# Print a model summary
summary(model)
```

Save our results

```{r save}
# Save the model
saveRDS(model, file = here::here("Models/Pass/", paste("Direct_Pass", strength, "model.Rda", sep = "_")))
```



