---
output: pdf_document
---

\centering
\Huge

**INLA SPDE Model - Entry Decision Model 1D**

\raggedright

\rule{\textwidth}{1.5pt}
\small

```{r setup, include = FALSE}
library(knitr) # For formatting the document
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

Summary: This file provides a template for determining the probability of a player opting for a control entry (carried or passed) upon entering the zone (vs dumping it in). Our model here uses logistic regression via INLA with a 1D spatial effect along the blue line.


# Load data and packages

Load necessary packages

```{r load packages}
library(tidyverse) # For various data analysis functions

library(here) # To call or save code/data in certain folders

library(INLA) # To fit model with INLA

source(here::here("Code/0. Functions/Plotting-Functions.R")) # For ggField2 function
```

Load data sets

```{r load data}
otters_data = read_csv(here::here("Data/otters_data_for_modelling.csv"))
```

# Specify Strength State

Specify which strength state we wish to run the model in

```{r parameters}
strength = "ES" # "PP" or "ES"
```


# Data Preparation

Filter our data down to what we wish to analyze (zone entries at the specified strength state).

```{r data prep}
# Filter down to our desired data
entry_data = otters_data %>%
  # Only zone entries
  filter(event == "Zone Entry") %>%
  # Only at the specified strength state
  filter(entry_strength == strength) %>%
  # Add in a binary variable for controlled (1) or dumped (0) that we will use as our response
  mutate(control = ifelse(detail_1 %in% c("Carried", "Played"), 1, 0))

# Extract a vector of y-coordinates from the model
model_coords = entry_data$same_direction_y
```


# Create Meshes

Create the 1D mesh as all possible coordinates along the blue line.

```{r mesh}
Mesh = inla.mesh.1d(0:85)
```


# Model Specifications

Set a prior and create an A matrix and w index for our model.

```{r spatial effect}
# Create a weight matrix
model_A_matrix = inla.mesh.1d.A(Mesh, loc = model_coords)

# Set the PC Matern prior
model_spde = inla.spde2.pcmatern(mesh = Mesh, prior.range = c(3, 0.8), prior.sigma = c(0.5, 0.5))

# Create spatial weight index
model_w_index = inla.spde.make.index("i", n.spde = model_spde$n.spde)
```

Create our model matrix of covariates (just score state and period).

```{r model matrix}
X = entry_data %>% select(offensive_score_diff, period)
```

Create a stack of all information to include in the model.

```{r stack}
model_stack = inla.stack(
  
  # Add our response
  data = list(y = entry_data$control), 
  
  # Add a list of the weights on the effects
  A = list(1, 1, model_A_matrix),
  
  # Add in effects
  effects = list(
    
    # Intercept, repeat 1 n times
    Intercept = rep(1, nrow(entry_data)),
    
    # Model matrix
    X = X,
    
    # Spatial effect
    model_w_index
  ))
```

Model formula

```{r formula}
# Create the formula
model_formula = as.formula(paste0("y ~ -1 + Intercept + period:offensive_score_diff + f(i, model = model_spde)"))
```


# Model Results

Run the model

```{r model}
model = inla(model_formula,
             
             family = "binomial",
             
             data = inla.stack.data(model_stack),
             
             control.compute = list(dic = TRUE, waic = TRUE, config = TRUE),
             
             control.predictor = list(A = inla.stack.A(model_stack)),
             
             num.threads = 8
)
```

Evaluate our model

```{r viz}
# Extract fitted values
fitval = model$summary.fitted.values[1:nrow(entry_data),] %>%
  mutate(y = entry_data$same_direction_y) %>%
  mutate(score_effect = entry_data$offensive_score_diff)

# Plot the mean probability of entry along the blue line
entry_1d = ggplot(fitval) +
  geom_point(aes(x = y, y = mean), alpha = 0.3) +
  labs(title = "1D Representation of Dump In to Recovery") +
  geom_smooth(aes(x = y, y = mean), colour = "green", size = 3) +
  labs(title = paste("Control Probability vs Location", strength, sep = " "), y = "Probability of Controlled Entry", x = "Rink Width")

# Plot the difference in entry probability for various score effects
entry_score_effect = ggplot(fitval) +
  geom_point(aes(x = score_effect, y = mean), alpha = 0.3) +
  labs(title = "1D Representation of Dump In to Recovery") +
  geom_smooth(aes(x = score_effect, y = mean), colour = "green", size = 3) +
  labs(title = "Control Prob vs Score Effect")

# Save the spatial map
ggsave(filename = paste("Entry Decision", strength, "1D INLA Mesh.png", sep = " "), plot = entry_1d, 
       device = "png", path = here::here("Figures/Entry/"))

# Print a model summary
summary(model)
```

Save our results

```{r save}
# Save the model
saveRDS(model, file = here::here("Models/Entry", paste("Entry_Decision", strength, "model.Rda", sep = "_")))
```



