---
output: pdf_document
---

\centering
\Huge

**INLA SPDE Model - One-Timer Probability Model with AR(1) Process for Time Since Last Pass**

\raggedright

\rule{\textwidth}{1.5pt}
\small

```{r setup, include = FALSE}
library(knitr) # For formatting the document
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

Summary: This file provides a template for building our spatiotemporal logistic regression model with INLA used in the simulations of our model. Here we are predicting the probability that a given shot is a one-timer.


# Load data and packages

Load necessary packages

```{r load packages}
library(tidyverse) # For various data analysis functions

library(here) # To call or save code/data in certain folders

library(INLA) # To fit model with INLA

source(here::here("Code/0. Functions/Plotting-Functions.R")) # For ggField2 function
```

Load data set

```{r load data}
otters_data = read_csv(here::here("Data/otters_data_for_modelling.csv"))
```

# Specify Strength State

Specify which strength state we wish to run the model in

```{r parameters}
strength = "ES" # "PP" or "ES"
```


# Data Preparation

Filter down to the data we wish to build the model with (shots that are preceded by a pass).

```{r data prep}
# Trim to our desired events (shots)
shot_data = otters_data %>%
  # Filter for shots
  filter(event %in% c("Shot", "Goal")) %>%
  # Filter for specified strength state
  filter(entry_strength == strength) %>%
  # Filter for shots that are preceded by a pass
  filter(is_prev_event_pass == 1) %>%
  # Convert logical variables to 0-1 binaries
  mutate(detail_3 = as.numeric(detail_3), 
         detail_4 = as.numeric(detail_4))
```


# Create Meshes

Build the spatial shot mesh.

```{r shot mesh}
# Create a coordinate matrix
locations = shot_data %>% 
  select(same_direction_x, same_direction_y) %>%
  mutate_all(as.numeric) %>%
  as.matrix()

# Set the boundary for the mesh
xvals = locations[,1]
yvals = locations[,2]

xrange = c(min(xvals), 189)
yrange = c(min(yvals), 85)

model_domain <- matrix(c(xrange[1], yrange[1],
                         xrange[2], yrange[1],
                         xrange[2], yrange[2],
                         xrange[1], yrange[2],
                         xrange[1], yrange[1]),
                       ncol = 2L, byrow = TRUE)


# Create the mesh
Mesh = inla.mesh.2d(locations, loc.domain = model_domain, max.edge = c(20, 15), offset = c(1, 2), cutoff = 8)


# Plot the mesh and data points
plot(Mesh)
points(locations, col = "blue", pch = 4)

# Save the mesh
saveRDS(Mesh, here::here(paste("Models/Meshes/One_Timer", strength, "Shot_Mesh.Rda", sep = "_")))
```

Build the spatial pass mesh.

```{r pass mesh}
# Create a coordinate matrix
pass_loc = shot_data %>% 
  select(last_pass_x, last_pass_y) %>%
  mutate_all(as.numeric) %>%
  as.matrix()

# Set the boundary for the mesh
pass_x = pass_loc[,1]
pass_y = pass_loc[,2]

pass_xrange = c(min(pass_x), 200)
pass_yrange = c(min(pass_y), 85)

pass_domain <- matrix(c(pass_xrange[1], pass_yrange[1],
                        pass_xrange[2], pass_yrange[1],
                        pass_xrange[2], pass_yrange[2],
                        pass_xrange[1], pass_yrange[2],
                        pass_xrange[1], pass_yrange[1]),
                      ncol = 2L, byrow = TRUE)

# Create the mesh
pass_Mesh = inla.mesh.2d(pass_loc, loc.domain = pass_domain, max.edge = c(20, 15), offset = c(1, 2), cutoff = 8)

# Plot the mesh
plot(pass_Mesh)
points(pass_loc, col = "red", pch = 4)

saveRDS(pass_Mesh, here::here(paste("Models/Meshes/One_Timer", strength, "Pass_Mesh.Rda", sep = "_")))
```

Build the temporal meshes. We create two here, one for shots built with random knots and one for passes with knots at 0, 1, 2, max seconds after the pass. We use random knots for shots since we did not want a temporal effect for it, but INLA must have a corresponding number of knots for all spatial effects.

```{r time mesh}
# Set the knots
knots = c(0, 1, 2, max(shot_data$time_since_last_pass))

knots_random = c(14, 27, 33, 58)

# Create 1D meshes
time_Mesh <- inla.mesh.1d(loc = knots)

random_Mesh = inla.mesh.1d(loc = knots_random)

# Save the meshes
saveRDS(time_Mesh, here::here(paste("Models/Meshes/One_Timer", strength, "Time_Mesh.Rda", sep = "_")))

saveRDS(random_Mesh, here::here(paste("Models/Meshes/One_Timer", strength, "Time_Mesh_Random.Rda", sep = "_")))
```

# Model Specifications

Set a prior and create a weight matrix and index for the shot spatial effect in our model.

```{r spatial shot}
# Create a spatial weight matrix for shot locations
model_A_matrix = inla.spde.make.A(Mesh, loc = locations, n.group = length(knots_random), group = shot_data$time_since_entry, group.mesh = random_Mesh)

# Set the PC Matern prior
model_spde = inla.spde2.pcmatern(mesh = Mesh, prior.range = c(30, 0.8), prior.sigma = c(0.5, 0.5))

# Create shot spatial weight index
model_w_index = inla.spde.make.index("w", n.spde = model_spde$n.spde, n.group = length(knots_random))
```

Set a prior and create a weight matrix and index for the pass spatial effect in our model.

```{r spatial pass}
# Create a spatial weight matrix for pass locations
pass_A_matrix = inla.spde.make.A(pass_Mesh, loc = pass_loc, n.group = length(knots), group = shot_data$time_since_last_pass, group.mesh = time_Mesh)

# Set the PC Matern prior
pass_spde = inla.spde2.pcmatern(mesh = pass_Mesh, prior.range = c(30, 0.8), prior.sigma = c(0.5, 0.5))

# Create pass spatial weight index
pass_w_index = inla.spde.make.index("g", n.spde = pass_spde$n.spde, n.group = length(knots))
```

Create a stack of all information to include in the model.

```{r stack}
# Initialize the stack
model_stack = inla.stack(
  
  # Add our response variable
  data = list(y = shot_data$detail_4), 
  
  # Add a list of the weights on the effects
  A = list(1, 1, pass_A_matrix, model_A_matrix),
  
  # Add spatial effects
  effects = list(
    
    # Intercept, repeat 1 n times
    Intercept = rep(1, nrow(shot_data)),
    
    # Add our model matrix (just use traffic as the only covariate)
    X = shot_data %>% select(detail_3),
    
    # Add pass spatial effect
    pass_w_index,
    
    # Add shot spatial effect
    model_w_index
    
  ))
```

State our model formula.

```{r formula}
model_formula = as.formula(paste0("y ~ -1 + Intercept ", 
                                  " + detail_3",
                                  " + f(w, model = model_spde, group = w.group, control.group = list(model = 'ar1'))",
                                  " + f(g, model = pass_spde, group = g.group, control.group = list(model = 'ar1'))"))
```


# Model Results

Run the model

```{r model}
model = inla(model_formula,
             
             family = "binomial",
             
             data = inla.stack.data(model_stack),
             
             control.compute = list(dic = TRUE, waic = TRUE, config = TRUE),
             
             control.predictor = list(A = inla.stack.A(model_stack)),
             
             num.threads = 8
)
```

Evaluate the model

```{r viz}
# Create the spatial maps
model_map_shot = gg_OZ(model, Mesh, Mat_Name = "w", Groups = 4, family = "binomial",
                  GroupLabels = c("14 Seconds Since Entry", "27 Seconds Since Entry", 
                                  "33 Seconds Since Entry", "58 Seconds Since Entry")) +
  labs(title = paste("Probability a Shot with a Pre-Shot Pass is a One-Timer ", strength, " AR(1) INLA Spatial Map", sep = ""))

model_map_pass = gg_OZ(model, pass_Mesh, Mat_Name = "g", Groups = 4, family = "binomial",
                       GroupLabels = c("0 Seconds Before Shot", "1 Second Before Shot", 
                                       "2 Seconds Before Shot", ">2 Seconds Before Shot")) +
  labs(title = paste("Pass Before One-Timer ", strength, " AR(1) INLA Spatial Map", sep = ""))

# Save the spatial maps
ggsave(filename = paste("One-Timer ", strength, " AR(1) INLA Spatial Map.png", sep = ""), 
       plot = model_map_shot, device = "png", path = here::here("Figures/Other/"))

ggsave(filename = paste("Pass Before One-Timer ", strength, " AR(1) INLA Spatial Map.png", sep = ""), 
       plot = model_map_pass, device = "png", path = here::here("Figures/Other/"))

# Print a model summary
summary(model)
```

Save our results

```{r save}
saveRDS(model, here::here(paste("Models/Other/One_Timer", strength, "model.Rda", sep = "_")))
```



