---
output: pdf_document
---

\centering
\Huge

**INLA SPDE Model - Pass Completion Model**

\raggedright

\rule{\textwidth}{1.5pt}
\small

```{r setup, include = FALSE}
library(knitr) # For formatting the document
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

Summary: This file provides a template for building our spatiotemporal logistic regression model with INLA to determine the probability that a pass is completed successfully.


# Load data and packages

Load necessary packages

```{r load packages}
library(tidyverse) # For various data analysis functions

library(here) # To call or save code/data in certain folders

library(INLA) # To fit model with INLA

library(raster, exclude = "select") # For distance calculations

library(LearnGeom) # For angle calculations

source(here::here("Code/0. Functions/Plotting-Functions.R")) # For ggField2 function
```

Load data set

```{r load data}
otters_data = read_csv(here::here("Data/otters_data_for_modelling.csv"))
```

# Specify Strength State

Specify which strength state we wish to run the model in

```{r parameters}
strength = "ES"
```


# Data Preparation

Filter for necessary data and add in extra features.

```{r data prep}
# Trim data and create additional variables
pass_data = otters_data %>%
  # Filter for passes
  filter(event %in% c("Pass", "Failed Pass")) %>%
  # Filter for only specified strength state
  filter(entry_strength == strength) %>%
  # Remove a few odd cases without pass target locations
  filter(!is.na(same_direction_x2)) %>%
  # Ignore if target is outside offensive zone
  filter(same_direction_x2 >= 125) %>%
  # Add variable for passes that come from below goal line
  mutate(from_behind_net = ifelse(last_pass_x > 189, 1, 0)) %>%
  # Determine if pass is complete
  mutate(is_complete = ifelse(event == "Pass",  1, 0)) %>%
  # Determine if pass is direct or not (1 = direct, 2 = indirect)
  mutate(is_direct = ifelse(detail_1 == "Direct", 1, 2)) %>%
  # Did the pass start behind the net?
  mutate(from_behind_net = ifelse(same_direction_x >= 189, 1, 0))

# Add in extra covariates for pass angle and distance
pass_data$angle = NA
pass_data$pass_distance = NA

for (i in 1:nrow(pass_data)) {
  
  pass_data$pass_distance[i] = pointDistance(c(pass_data$same_direction_x[i], pass_data$same_direction_y[i]), 
                                             c(pass_data$same_direction_x2[i], pass_data$same_direction_y2[i]),
                                             lonlat = FALSE)
  
  if (!((pass_data$same_direction_x[i] == pass_data$same_direction_x2[i]) & (pass_data$same_direction_y[i] == pass_data$same_direction_y2[i]))) {
    pass_data$angle[i] = Angle(A = c(pass_data$same_direction_x[i], 95), 
                             B = c(pass_data$same_direction_x[i], pass_data$same_direction_y[i]),
                             C = c(pass_data$same_direction_x2[i], pass_data$same_direction_y2[i]))
  } else {pass_data$angle[i] = 0}
  
  pass_data$angle[i] = case_when(
    (pass_data$same_direction_x[i] > pass_data$same_direction_x2[i]) & (pass_data$same_direction_y[i] > pass_data$same_direction_y2[i]) ~ abs(pass_data$angle[i] - 180),
    (pass_data$same_direction_x[i] > pass_data$same_direction_x2[i]) & (pass_data$same_direction_y[i] <= pass_data$same_direction_y2[i]) ~ abs(-pass_data$angle[i]),
    (pass_data$same_direction_x[i] <= pass_data$same_direction_x2[i]) & (pass_data$same_direction_y[i] > pass_data$same_direction_y2[i]) ~ abs(180 - pass_data$angle[i]),
    (pass_data$same_direction_x[i] <= pass_data$same_direction_x2[i]) & (pass_data$same_direction_y[i] <= pass_data$same_direction_y2[i]) ~ abs(pass_data$angle[i]),
    is.nan(pass_data$angle[i]) ~ 0
  )
}

# Determine if the pass is moving forwards or backwards
pass_data = pass_data %>%
  mutate(is_forwards = ifelse(same_direction_x <= same_direction_x2, 1, 0)) %>%
  mutate(is_backwards = ifelse(same_direction_x > same_direction_x2, 1, 0))
```



# Create Meshes

Build the spatial pass start location mesh.

```{r shot mesh}
# Create a coordinate matrix
start_loc = pass_data %>% 
  select(same_direction_x, same_direction_y) %>%
  mutate_all(as.numeric) %>%
  as.matrix()

# Set the boundary for the mesh
xvals = start_loc[,1]
yvals = start_loc[,2]

xrange = c(min(xvals), 189)
yrange = c(min(yvals), 85)

model_domain <- matrix(c(xrange[1], yrange[1],
                         xrange[2], yrange[1],
                         xrange[2], yrange[2],
                         xrange[1], yrange[2],
                         xrange[1], yrange[1]),
                       ncol = 2L, byrow = TRUE)

# Create the mesh
start_Mesh = inla.mesh.2d(start_loc, loc.domain = model_domain, max.edge = c(20, 15), offset = c(1, 2), cutoff = 8)

# Plot the mesh and data points
plot(start_Mesh)
points(start_loc, col = "blue", pch = 4)

saveRDS(start_Mesh, here::here(paste("Models/Meshes/Pass_Completion_Start", strength, "Mesh.Rda", sep = "_")))
```

Build the spatial pass end location mesh.

```{r pass mesh}
# Create a coordinate matrix
end_loc = pass_data %>% 
  select(same_direction_x2, same_direction_y2) %>%
  mutate_all(as.numeric) %>%
  as.matrix()

# Set the boundary for the mesh
pass_x = end_loc[,1]
pass_y = end_loc[,2]

pass_xrange = c(min(pass_x), 189)
pass_yrange = c(min(pass_y), 85)

pass_domain <- matrix(c(pass_xrange[1], pass_yrange[1],
                        pass_xrange[2], pass_yrange[1],
                        pass_xrange[2], pass_yrange[2],
                        pass_xrange[1], pass_yrange[2],
                        pass_xrange[1], pass_yrange[1]),
                      ncol = 2L, byrow = TRUE)

# Create the mesh
end_Mesh = inla.mesh.2d(end_loc, loc.domain = pass_domain, max.edge = c(20, 15), offset = c(1, 2), cutoff = 8)

# Plot the mesh
plot(end_Mesh)
points(end_loc, col = "red", pch = 4)

saveRDS(end_Mesh, here::here(paste("Models/Meshes/Pass_Completion_End", strength, "Mesh.Rda", sep = "_")))
```




# Model Specifications

Set a prior and create a weight matrix and index for the pass start spatial effect in our model.

```{r shot spatial}
# Create a weight matrix
start_A_matrix = inla.spde.make.A(start_Mesh, loc = start_loc, n.group = 2, group = pass_data$is_direct)

# Set the PC Matern prior
start_spde = inla.spde2.pcmatern(mesh = start_Mesh, prior.range = c(35, 0.5), prior.sigma = c(0.5, 0.5))

# Create weight index
start_w_index = inla.spde.make.index("w", n.spde = start_spde$n.spde, n.group = 2)
```

Set a prior and create a weight matrix and index for the pass end spatial effect in our model.

```{r pass spatial}
# Create a weight matrix
end_A_matrix = inla.spde.make.A(end_Mesh, loc = end_loc, n.group = 2, group = pass_data$is_direct)

# Set the PC Matern prior
end_spde = inla.spde2.pcmatern(mesh = end_Mesh, prior.range = c(40, 0.7), prior.sigma = c(0.5, 0.5))

# Create weight index
end_w_index = inla.spde.make.index("g", n.spde = end_spde$n.spde, n.group = 2)
```

Create our model matrix.

```{r model matrix}
X = pass_data %>%
  # Select necessary columns
  select(pass_time_decay, pass_count, pass_distance, 
         angle, is_forwards, is_backwards, from_behind_net)
```

Create a stack of all information to include in the model.

```{r stack}
# Initialize the stack
model_stack = inla.stack(
  
  # Add our response variable
  data = list(y = pass_data$is_complete), 
  
  # Add a list of the weights on the effects
  A = list(1, 1, start_A_matrix, end_A_matrix),
  
  # Add spatial effects
  effects = list(
    
    # Intercept, repeat 1 n times
    Intercept = rep(1, nrow(pass_data)),
    
    # Add our model matrix
    X = X,
    
    # Add pass start spatial effect
    start_w_index,
    
    # Add pass end spatial effect
    end_w_index
    
  ))
```

State our model formula.

```{r formula}
model_formula = as.formula(paste0("y ~ -1 + Intercept ", 
                                  " + pass_time_decay + pass_distance + angle:is_forwards + angle:is_backwards",
                                  " + pass_count + from_behind_net",
                                  " + f(w, model = start_spde, group = w.group, control.group = list(model = 'exchangeable'))",
                                  " + f(g, model = end_spde, group = g.group, control.group = list(model = 'exchangeable'))"))
```


# Model Results

Run the model

```{r model}
model = inla(model_formula, 
             
             family = "binomial",
             
             data = inla.stack.data(model_stack),
             
             control.compute = list(dic = TRUE, waic = TRUE, config = TRUE),
             
             control.predictor = list(A = inla.stack.A(model_stack)),
             
             num.threads = 8
)
```

Evaluate our model

```{r viz}
# Create the spatial maps
model_map_start = gg_OZ(model, start_Mesh, Mat_Name = "w", Groups = 2, family = "binomial",
                  GroupLabels = c("Direct Pass", "Indirect Pass")) +
  labs(title = paste("Pass Completion Start ", strength, " AR(1) INLA Spatial Map", sep = ""))

model_map_end = gg_OZ(model, end_Mesh, Mat_Name = "g", Groups = 2, family = "binomial",
                  GroupLabels = c("Direct Pass", "Indirect Pass")) +
  labs(title = paste("Pass Completion End ", strength, " AR(1) INLA Spatial Map", sep = ""))

# Save the spatial maps
ggsave(filename = paste("Pass Completion Start ", strength, " AR(1) INLA Spatial Map.png", sep = ""), 
       plot = model_map_start, device = "png", path = here::here("Figures/Pass/"))

ggsave(filename = paste("Pass Completion End ", strength, " AR(1) INLA Spatial Map.png", sep = ""), 
       plot = model_map_end, device = "png", path = here::here("Figures/Pass/"))

# Print a model summary
summary(model)
```

Save our results

```{r save}
# Save the model
saveRDS(model, file = here::here("Models/Pass/", paste("Pass_Completion", strength, "model.Rda", sep = "_")))
```



