---
output: pdf_document
---

\centering
\Huge

**INLA SPDE Model - Time Elapsed Model**

\raggedright

\rule{\textwidth}{1.5pt}
\small

```{r setup, include = FALSE}
library(knitr) # For formatting the document
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

Summary: This file provides a template for building our spatiotemporal regression model with INLA used in the simulations of our model. Here we are predicting the time ellapsed before another event occurs.


# Load data and packages

Load necessary packages

```{r load packages}
library(tidyverse) # For various data analysis functions

library(here) # To call or save code/data in certain folders

library(INLA) # To fit model with INLA

source(here::here("Code/0. Functions/Plotting-Functions.R")) # For ggField2 function
```

Load data set

```{r load data}
otters_data = read_csv(here::here("Data/otters_data_for_modelling.csv"))
```

# Specify Strength State

Specify which strength state we wish to run the model in

```{r parameters}
strength = "ES" # "PP" or "ES"
```


# Data Preparation

Filter down to the data we wish to build the model with and add in other useful variables.

```{r data prep}
# Add necessary variables
otters_data = otters_data %>%
  # Filter for strength state
  filter(entry_strength == strength) %>%
  # Clean event names
  mutate(event = ifelse(event == "Failed Pass", "Pass", event)) %>%
  mutate(event = ifelse(event == "Goal", "Shot", event)) %>%
  # Turn direct/indirect pass into a binary
  mutate(is_direct = ifelse(detail_1 == "Direct", 1, 0) %>% as.numeric()) %>%
  # Was the previous event a direct pass?
  mutate(prev_direct_pass = ifelse(prev_event == "Pass" & detail_1 == "Direct", 1, 0)) %>%
  # Was the previous event an indirect pass?
  mutate(prev_indirect_pass = ifelse(prev_event == "Pass" & detail_1 == "Indirect", 1, 0)) %>%
  # Previous event binaries
  mutate(prev_shot = ifelse(prev_event == "Shot", 1, 0)) %>%
  mutate(prev_recovery = ifelse(prev_event == "Recovery", 1, 0)) %>%
  mutate(prev_turnover = ifelse(prev_event == "Turnover", 1, 0)) %>%
  # Nest by entry-to-exit sequence
  group_by(entry_id) %>%
  nest() %>%
  # For each entry-to-exit sequence...
  mutate(data = map(data, ~.x %>%
                      # Get the previous xy-coordinates for all events and time ellapsed
                      mutate(prev_x = lag(same_direction_x),
                             prev_y = lag(same_direction_y),
                             time_since_prev_event = seconds - lag(seconds)))) %>%
  unnest() %>% ungroup() %>%
  # Remove entries
  filter(event != "Zone Entry") %>%
  # Determine distance from last event
  mutate(distance_last_event = (same_direction_x - prev_x) ^ 2 + (same_direction_y - prev_y) ^ 2) %>%
  # Create interaction terms
  mutate(int_direct_dist = distance_last_event * prev_direct_pass) %>%
  mutate(int_indirect_dist = distance_last_event * prev_indirect_pass) %>%
  mutate(int_turnover_dist = distance_last_event * prev_turnover) %>%
  mutate(int_recovery_dist = distance_last_event * prev_recovery) %>%
  mutate(int_shot_dist = distance_last_event * prev_shot) %>%
  # Label events as numeric
  mutate(type_index = plyr::mapvalues(event, from = unique(.$event), to = c(1:length(unique(.$event)))) %>% as.numeric())
```


# Create Meshes

Build the spatial event mesh.

```{r shot mesh}
# Create the boundary for OZ mesh
oz_boundary = data.frame(
  x = c(125, 200, 200, 125, 125),
  y = c(0, 0, 85, 85, 0)
)

model_domain = as.matrix(oz_boundary)


# Get model coordinates and matrix
model_coords = as.matrix(otters_data %>% select(same_direction_x, same_direction_y))
model_matrix = otters_data %>% select(same_direction_x, same_direction_y, is_direct)


# Build INLA SPDE mesh
Mesh = inla.mesh.2d(model_coords, max.edge = c(20, 15), offset = c(1, 2), cutoff = 8, loc.domain = model_domain)


# Plot the mesh and data points
plot(Mesh)
points(model_coords, col = "blue", pch = 4)


# Save the mesh
saveRDS(Mesh, here::here(paste("Models/Meshes/Time_to_Event", strength, "Mesh.Rda", sep = "_")))
```


# Model Specifications

Set a prior and create a weight matrix and index for the spatial effect in our model.

```{r spatial shot}
# Create a spatial weight matrix for shot locations
model_A_matrix = inla.spde.make.A(Mesh, loc = model_coords, n.group = 5, group = otters_data$type_index)

# Set the PC Matern prior
model_spde = inla.spde2.pcmatern(mesh = Mesh, prior.range = c(5, 0.8), prior.sigma = c(0.5, 0.5))

# Create shot spatial weight index
model_w_index = inla.spde.make.index("w", n.spde = model_spde$n.spde, n.group = 5)
```

Create model matrix of covariates.

```{r model matrix}
X = otters_data %>%
  select(prev_turnover, prev_recovery, prev_shot, distance_last_event, int_direct_dist, prev_indirect_pass,
         int_indirect_dist, int_turnover_dist, int_recovery_dist, int_shot_dist)
```


# Model Specifications

Create a stack of all information to include in the model.

```{r stack}
# Initialize the stack
model_stack = inla.stack(
  
  # Add our response variable
  data = list(y = otters_data$time_since_prev_event), 
  
  # Add a list of the weights on the effects
  A = list(1, 1, model_A_matrix),
  
  # Add spatial effects
  effects = list(
    
    # Intercept, repeat 1 n times
    Intercept = rep(1, nrow(otters_data)),
    
    # Add covariates
    X = X,
    
    # Add shot spatial effect
    model_w_index
    
  ))
```

State our model formula.

```{r formula}
model_formula = as.formula(paste0("y ~ -1 + Intercept ", 
                                  " + prev_indirect_pass + prev_turnover + prev_recovery + prev_shot",
                                  " + distance_last_event + int_direct_dist + int_indirect_dist + int_turnover_dist",
                                  " + int_recovery_dist + int_shot_dist",
                                  " + f(w, model = model_spde, group = w.group, control.group = list(model = 'iid'))"))
```

# Model Results

Run the model

```{r model}
model = inla(model_formula,
             
             family = "gaussian",
             
             data = inla.stack.data(model_stack),
             
             control.compute = list(dic = TRUE, waic = TRUE, config = TRUE),
             
             control.predictor = list(A = inla.stack.A(model_stack)),
             
             num.threads = 16
)
```


Evaluate the model

```{r viz}
# Get order of events
events = otters_data %>% select(event) %>% unlist() %>% unique()
labels = otters_data %>% select(type_index) %>% unlist() %>% unique()

# Create the spatial maps
model_map = gg_OZ(model, Mesh, Mat_Name = "w", Groups = 4, 
                  GroupLabels = events) +
  labs(title = paste("Time to Event ", strength, " IID INLA Spatial Map", sep = "")) +
  scale_fill_viridis_d()

# Save the spatial maps
ggsave(filename = paste("Time to Event ", strength, " IID INLA Spatial Map.png", sep = ""), 
       plot = model_map, device = "png", path = here::here("Figures/Other/"))

# Print a model summary
summary(model)
```

Save our results

```{r save}
saveRDS(model, here::here(paste("Models/Other/Time_to_Event", strength, "model.Rda", sep = "_")))
```

